---
title: "Bat phylogenetic tree"
author: "Lucas Nell"
date: "`r format(Sys.Date())`"
output: github_document
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(comment = '', prompt = FALSE)
```


# Background

The paper by 
[Agnarsson et al. (2011)](http://dx.doi.org/10.1371/currents.RRN1212) used only one
mitochrondrial DNA gene and is the reason the Open Tree of Life (OTOL) version doesn't 
have a single resolved branch for *Myotis lucifugus*. A more recent paper by 
[Shi and Rabosky (2015)](http://onlinelibrary.wiley.com/doi/10.1111/evo.12681/abstract) 
uses multiple mitochondrial and nuclear genes and contains *M. lucifugus*.
This paper will be used for this tree.


# Preamble

I first need to `source` the `'tree_preamble.R'` file to load necessary R packages and 
make the data frame of species names.

```{r, message = FALSE, warning = FALSE}
source('tree_preamble.R')
```


# Retrieving tree

You can download the tree by 
[Shi and Rabosky (2015)](http://onlinelibrary.wiley.com/doi/10.1111/evo.12681/abstract)
(OTOL link [here](https://tree.opentreeoflife.org/curator/study/view/ot_254)),
adjust its tip labels to remove underscores,
and drop outgroups as such:

```{r}
bat_tr <- get_study(study_id = 'ot_254', object_format = 'phylo')
bat_tr$tip.label <- gsub('_', ' ', bat_tr$tip.label)
bat_tr <- drop.tip(bat_tr, tip = c('Mus musculus', 'Sorex araneus', 'Canis lupus'))
bat_tr
```



# Checking species names

All bat species are present in this tree.

```{r}
sp_df$species[sp_df$type == 'bat' & ! sp_df$species %in% bat_tr$tip.label]
```


# Drawing tree

If you want to make a pdf of this tree that's (somewhat) readable:

```{r, eval = FALSE}
pdf('bat_tree.pdf', width = 6, height = 24)
    plot(bat_tr, cex = 0.1, edge.width = 0.25, no.margin = TRUE, label.offset = 0.5)
    nodelabels(frame = 'none', cex = 0.1, adj = c(1, 0))
    tiplabels(frame = 'none', cex = 0.1, adj = c(0, 0.5))
dev.off()
```



