---
title: "Phylogenetic tree"
author: "Lucas Nell"
output:
  html_document:
    highlight: pygments
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = TRUE, comment = '', prompt = FALSE)
```


# Background

The paper by 
[Agnarsson et al. (2011)](http://dx.doi.org/10.1371/currents.RRN1212) used only one
mitochrondrial DNA gene and is the reason the Open Tree of Life (OTOL) version doesn't 
have a single resolved branch for *Myotis lucifugus*. A more recent paper by 
[Shi and Rabosky (2015)](http://onlinelibrary.wiley.com/doi/10.1111/evo.12681/abstract) 
uses multiple mitochondrial and nuclear genes and contains *M. lucifugus*.
This paper will be used to complete the tree.


# Attempting OTOL tree

I used the package `rotl` to interface with the Open Tree of Life (OTOL) to import trees
into R. [Here](https://cran.r-project.org/web/packages/rotl/vignettes/data_mashups.html)
is a guide on using it.

```{r, message = FALSE, warning = FALSE}
library(dplyr)
library(readr)
library(tidyr)
library(ggplot2)
library(ape)
library(rotl)
# library(phylolm)
# library(vegan)
# library(lme4)
# library(pez)
# library(lmerTest)
```


Data frame of species needed:

```{r}
sp_df <- 
'species,type
Mus musculus,rodent
Rattus norvegicus,rodent
Thaptomys nigrita,rodent
Akodon montensis,rodent
Delomys sublineatus,rodent
Olygoryzomys nigripes,rodent
Sooretamys angouya,rodent
Euryoryzomys russatus,rodent
Peromyscus leucopus,rodent
Microtus pennsylvanicus,rodent
Artibeus lituratus,bat
Carollia perspicillata,bat
Desmodus rotundus,bat
Molossus rufus,bat
Molossus molossus,bat
Eumops glaucinus,bat
Tadarida brasiliensis,bat
Myotis lucifugus,bat
Eptesicus fuscus,bat
' %>% 
    read_csv(.)
```


Searching for these species names in the OTOL database to make sure they exist there and 
to get their species IDs.

```{r}
taxon_search <- tnrs_match_names(names = sp_df$species, context_name = 'Mammals')
knitr::kable(taxon_search)
```


Making sure our names and IDs match up with those of the OTOL.

```{r}
sp_df$ott_name <- taxon_search$unique_name
sp_df$ott_id <- taxon_search$ott_id
```


We don't have good representation in published trees from this search...

```{r}
hits <- lapply(sp_df$ott_id, studies_find_trees, property="ot:ottId", detailed = FALSE)
sapply(hits, function(x) sum(x[["n_matched_trees"]]))
```

So I'm creating a tree from the synthesis OTOL tree using species, except replacing 
*M. lucifugus* with *M. lucifugus lucifugus*. Using *M. lucifugus lucifugus* allows me to
see where *M. lucifugus* will probably be located.

> This tree only serves to guide future searches, because the OTOL tree has no branch 
> lengths, so cannot be used to generate a phylogenetic covariance matrix.

```{r}
tr <- tol_induced_subtree(
    ott_ids = c(sp_df$ott_id[sp_df$species != 'Myotis lucifugus'],
                tnrs_match_names(names = 'Myotis lucifugus lucifugus', 
                                 context_name = 'Mammals')$ott_id))

tr$tip.label <- strip_ott_ids(tr$tip.label, remove_underscores=TRUE)
plot(tr, no.margin = TRUE, label.offset = 0.5); nodelabels(); tiplabels()
```


From more extensive searching on the OTOL database online, I found one tree that has all 
focal rodent species
[(Fabre et al. 2012)](http://dx.doi.org/10.1186/1471-2148-12-88)
and another that has all bat species
[(Shi and Rabosky 2015)](http://onlinelibrary.wiley.com/doi/10.1111/evo.12681/abstract).



# Rodent tree

Using `rotl` directly didn't work for this paper's tree, so I downloaded the Newick 
version directly from the link on 
[OTOL's site](https://tree.opentreeoflife.org/curator/study/view/pg_2688/?tab=metadata).

```{r, engine = 'bash', eval = FALSE}
wget "https://api.opentreeoflife.org/v3/study/pg_2688.tre"
mv pg_2688.tre rodents.tre
```

I then opened this file in a text editor and removed all single quotes and replaced all
spaces with underscores. This allows R to read this file appropriately. After it's read,
I can replace underscores with spaces.

```{r}
rod_tr <- read.tree('rodents.tre')
rod_tr$tip.label <- gsub('_', ' ', rod_tr$tip.label)
rod_tr$node.label <- gsub('_', ' ', rod_tr$node.label)
```


Three species don't seem to be present in either node or tip labels.

```{r}
sp_df$species[sp_df$type == 'rodent' & 
                  !sp_df$species %in% c(rod_tr$node.label, rod_tr$tip.label)]
```


Looking these up in the OTOL database, I found the following:
- *Olygoryzomys nigripes* is spelled *Oligoryzomys nigripes*
  [link](https://tree.opentreeoflife.org/taxonomy/browse?id=752853)
- *Sooretamys angouya* is synonymous with *Oryzomys angouya*
  [link](https://tree.opentreeoflife.org/taxonomy/browse?id=1039661)
- *Euryoryzomys russatus* is synonymous with *Oryzomys russatus*
  [link](https://tree.opentreeoflife.org/taxonomy/browse?id=739)

When these are taken into account, all species are found in this tree.

```{r}

```


# Bat tree

You can download the tree by 
[Shi and Rabosky (2015)](http://onlinelibrary.wiley.com/doi/10.1111/evo.12681/abstract)
[(OTOL link)](https://tree.opentreeoflife.org/curator/study/view/ot_254)
and adjust its tip labels to remove underscores as such:

```{r}
bat_tr <- get_study(study_id = 'ot_254', object_format = 'phylo')
bat_tr$tip.label <- gsub('_', ' ', bat_tr$tip.label)
```


If you want to make a pdf of this tree that's somewhat readable:
```{r, eval = FALSE}
pdf('Shi_tree.pdf', width = 6, height = 12)
    plot(bat_tr, cex = 0.1, edge.width = 0.25, no.margin = TRUE, label.offset = 0.5)
    nodelabels(frame = 'none', cex = 0.1, adj = c(1, 0))
    tiplabels(frame = 'none', cex = 0.1, adj = c(0, 0.5))
dev.off()
```


```{r}
# Rooting tree
# root(bat_tr, )
# Removing all tips except those for *Eptesicus fuscus* and *Myotis lucifugus*
bat_tr_filt <- drop.tip(
    bat_tr, 
    tip = bat_tr$tip.label[!bat_tr$tip.label %in% c('Eptesicus fuscus', 'Myotis lucifugus', 
                                              'Mus musculus')])
# bat_tr_filt <- root(bat_tr_filt, 'Mus musculus')

plot(bat_tr_filt); nodelabels(); tiplabels()
```


```{r}
is.rooted(drop.tip(
    bat_tr, 
    tip = bat_tr$tip.label[!bat_tr$tip.label %in% c('Mus musculus', 'Eptesicus fuscus', 
                                              'Myotis lucifugus')]))
nodelabels(); tiplabels()
```



```{r}
# Combining them
# ?bind.tree

# final_tr <- bind.tree(tr, bat_tr_filt, where = 23)
# plot(final_tr)
```



Current plan:
1. Remove bats from `bat_tr` that are present in `tr`
2. Combine them
3. Remove all species not in `sp_df$species`
