---
title: "Test whether to include log_mass in phylogenetic linear regression"
author: "Lucas Nell"
date: "`r Sys.setenv(TZ='America/Chicago'); format(Sys.Date(), '%d %b %Y')`"
output:
  github_document:
    toc: true
    toc_depth: 2
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE, cache = FALSE, purl = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
# Change this to TRUE if you want to run the models instead of getting them from 
# RDS files:
run_models <- FALSE
```


This file determines whether to use `log_mass` in `phylolm` regressions and 
`cor_phylo` correlations.
These analyses take about half an hour to run.

All p-values below are for whether the coefficient for log(mass) are not equal to zero.


```{r load_packages, echo = FALSE}
suppressPackageStartupMessages({
    library(readr)
    library(dplyr)
    library(tidyr)
    library(purrr)
    library(phylolm)
    library(ape)
    library(phyr)
})
invisible(sapply(list.files('R', '*.R', full.names = TRUE), source))
```
```{r load_models, echo = FALSE, eval = !run_models}
diet_fit <- read_rds('output/inc_mass_diet_fit.rds')
absorp_fit <- read_rds('output/inc_mass_absorp_fit.rds')
spp_fits <- read_rds('output/inc_mass_spp_fits.rds')
pos_fits <- read_rds('output/inc_mass_pos_fits.rds')
clear_sef <- read_rds('output/inc_mass_cor_phylo.rds')
```


# Necessary data:

```{r absorp_data}
# Morphometrics by species
spp_df <- get_df('spp')
tr <- get_tr('spp')
# Morphometrics by species and segment
seg_types <- c('prox', 'med', 'dist')
pos_df_list <- lapply(seg_types, get_df, .df = 'pos')
names(pos_df_list) <- seg_types
# Absorption by species
absorp_df <- get_df('absorp')
absorp_se_df <- get_df('absorp', .stat = 'se')  # <-- contains standard errors
absorp_tr <- get_tr('absorp')
# Clearance by species
clear_df <- get_df('clear')
clear_se_df <- get_df('clear', .stat = 'se')  # <-- contains standard errors
clear_tr <- get_tr('clear')
```




# `SEF` on `Diet`

`phylolm` call and output:

```{r sef_diet, eval = run_models}
set.seed(29851644)
diet_fit <- phylolm(log_sef ~ diet + log_mass, data = spp_df, phy = tr,
                    model = 'lambda', boot = 2000)
```
```{r write_diet, echo = FALSE, eval = run_models}
write_rds(diet_fit, 'output/inc_mass_diet_fit.rds')
```

P-value:

```{r sef_diet_pval, echo = FALSE}
cat(sprintf('P = %.3f\n', pval(diet_fit, 'log_mass')))
```


# `Absorption` on `Clade`


`phylolm` call and output:

```{r absorp_clade, eval = run_models}
set.seed(1092141389)
absorp_fit <- suppressWarnings(  # gives warning about lambda being very low
    phylolm(log_absorp ~ clade + log_mass, data = absorp_df, phy = absorp_tr, 
            model = 'lambda', boot = 2000)
)
```
```{r write_absorp, echo = FALSE, eval = run_models}
write_rds(absorp_fit, 'output/inc_mass_absorp_fit.rds')
```

P-value:

```{r absorp_clade_pval, echo = FALSE}
cat(sprintf('P = %.3f\n', pval(absorp_fit, 'log_mass')))
```



# `Morphometrics` on `Clade`

`phylolm` calls:

```{r sp_analyses, eval = run_models}
spp_ys <- c("log_intestinal_length", "log_nsa", "log_vill_surface_area",
            "log_total_enterocytes")
set.seed(357885189)
spp_fits <- lapply(
    spp_ys,
    function(y) {
        f <- paste(y, '~ clade + log_mass')
        suppressWarnings(
            do.call("phylolm", list(as.formula(f), data = as.name("spp_df"),
                                    phy = as.name("tr"), model = 'lambda',
                                    boot = 2000))
        )
    })
names(spp_fits) <- spp_ys
```
```{r write_spp, echo = FALSE, eval = run_models}
write_rds(spp_fits, 'output/inc_mass_spp_fits.rds')
```

P-values:

```{r sp_analyses_pvals, echo = FALSE}
cbind(P = sapply(spp_fits, pval, 'log_mass'))
```





# `Morphometrics` on `Clade`, separately by segment


`phylolm` calls:


```{r pos_analyses, eval = run_models}
pos_ys <- c('log_intestinal_diameter', 'log_villus_height', 'villus_width', 
            'crypt_width', 'log_sef', 'enterocyte_diameter', 'log_enterocyte_density')
set.seed(632929430)
pos_fits <- lapply(
    seg_types,
    function(pos) {
        # Assigning to obj named <pos>_df so that the call identifies the position
        assign(paste0(pos, '_df'), pos_df_list[[pos]])
        lapply(
            pos_ys,
            function(y) {
                f <- paste(y, ' ~ clade + log_mass')
                arg_list <- list(
                    formula = as.formula(f),
                    data = as.name(paste0(pos, "_df")),
                    phy = as.name("tr"), model = 'lambda')
                # Some models don't find the peak likelihood unless specifying a
                # starting value of 0.1.
                LL_nostart <- suppressWarnings(do.call("phylolm", arg_list))$logLik
                arg_list_wstart <- c(arg_list, starting.value = quote(list(lambda = 0.1)))
                LL_wstart <- suppressWarnings(do.call("phylolm", arg_list_wstart))$logLik
                if (LL_wstart > LL_nostart) arg_list <- arg_list_wstart
                # Adding number of bootstrap replicates
                arg_list <- c(arg_list, boot = 2000)
                # Now call phylolm
                suppressWarnings(do.call("phylolm", arg_list))
            })
    })
names(pos_fits) <- seg_types
for (i in 1:length(pos_fits)) names(pos_fits[[i]]) <- pos_ys; rm(i)
```
```{r write_pos, echo = FALSE, eval = run_models}
write_rds(pos_fits, 'output/inc_mass_pos_fits.rds')
```


Because there are so many models, I'm writing whether to include mass to a csv file:

```{r csv_pos}
include_df <- lapply(
    seg_types, 
    function(p) {
        data_frame(pos = p, y = names(pos_fits[[p]]),
                   include = sapply(pos_fits[[p]], pval, 
                                    params = 'log_mass') < 0.05)
    }) %>% 
    bind_rows
write_csv(include_df, 'output/include_mass_pos.csv')
```




P-values:

```{r pos_analyses_pvals, echo = FALSE}
cat("proximal:\n")
cbind(P = sapply(pos_fits$prox, pval, params = 'log_mass'))
cat("medial:\n")
cbind(P = sapply(pos_fits$med, pval, params = 'log_mass'))
cat("distal:\n")
cbind(P = sapply(pos_fits$dist, pval, params = 'log_mass'))
```








# `Clearance` and `SEF`

In `cor_phylo`, you can include covariate(s) that might be having confounding effects.

Below, I'm trying out whether either clearance or SEF needs body mass as a covariate
by including it separately for clearance, then for SEF.



```{r clear_sef, eval = run_models}
# Making data frame with all info, including measurement errors:
clear_df_bm <- clear_df
clear_df_bm$log_clear_se <- clear_se_df$log_clear
clear_df_bm$log_sef_se <- clear_se_df$log_sef
# For this comparison, I have to remove one row that doesn't have body mass
clear_df_bm <- clear_df_bm[!is.na(clear_df$log_mass),]

# Phylogenetic tree, removing species with no body mass data
clear_tr_bm <- ape::drop.tip(
    clear_tr,
    tip = clear_tr$tip.label[!clear_tr$tip.label %in% clear_df_bm$species])

# Options that are the same for both calls
opts <- list(traits = quote(list(log_clear, log_sef)),
             meas_errors = quote(list(log_clear_se, log_sef_se)),
             species = quote(species), data = quote(clear_df_bm),
             phy = quote(clear_tr_bm), boot = 2000,
             method = "nelder-mead-r",
             constrain_d = TRUE,
             max_iter = 1e6, keep_boots = "fail")
# Now separating into different lists:
opts <- list(clear = c(opts, covariates = quote(list(log_clear = log_mass))),
             sef = c(opts, covariates = quote(list(log_sef = log_mass))))

# # cor_phylo run with bootstrapping (takes ~1.5 min)
# set.seed(1844365955)
# clear_sef <- map(c("clear", "sef"), ~ do.call(cor_phylo, opts[[.x]])) %>% 
#     setNames(c("clear", "sef"))
# # readr::write_rds(clear_sef, "~/Desktop/clear_sef.rds")
clear_sef <- readr::read_rds("~/Desktop/clear_sef.rds")
```


```{r cor_phylo_refits}

source(textConnection(readLines("reports/03-include_mass.Rmd")[30:39]))
source(textConnection(readLines("reports/03-include_mass.Rmd")[53:67]))
source(textConnection(readLines("reports/03-include_mass.Rmd")[239:268]))






cp_obj <- clear_sef$clear

new_call <- cp_obj$call
new_call$boot <- NULL
new_call$keep_boots <- NULL

data <- eval(new_call$data)
phy <- phyr:::check_phy(eval(new_call$phy))

# If a character, convert to an expression before evaluating in cp_get_species():
spp_vec <- phyr:::cp_get_species(new_call$species, data, phy)

phy_order <- match(phy$tip.label, spp_vec)
X <- phyr:::extract_traits(new_call$traits, phy_order, data)
trait_names <- colnames(X)
U <- phyr:::call_arg(new_call, "covariates")
U <- phyr:::extract_covariates(U, phy_order, trait_names, data)
M <- phyr:::call_arg(new_call, "meas_errors")
M <- phyr:::extract_meas_errors(M, phy_order, trait_names, data)

species <- phy$tip.label

new_call$traits <- quote(X)
new_call$species <- quote(species)
new_call$phy <- quote(phy)
new_call$covariates <- quote(U)
new_call$meas_errors <- quote(M)
new_call$data <- NULL

# new_args <- list(...)
new_args <- list(rcond_threshold = 1e-3)
for (x in names(new_args)) {
    new_call[[x]] <- new_args[[x]]
}

new_cps <- as.list(rep(NA, length(cp_obj$bootstrap$inds)))

for (i in 1:length(cp_obj$bootstrap$inds)) {
    
    for (j in 1:ncol(X)) {
        X[,j] <- cp_obj$bootstrap$mats[[i]][,j]
    }
    new_cps[[i]] <- eval(new_call)
    
    new_cps[[i]]$call$traits <- cp_obj$call$traits
    new_cps[[i]]$call$species <- cp_obj$call$species
    new_cps[[i]]$call$phy <- cp_obj$call$phy
    new_cps[[i]]$call$covariates <- cp_obj$call$covariates
    new_cps[[i]]$call$meas_errors <- cp_obj$call$meas_errors
    new_cps[[i]]$call$data <- cp_obj$call$data
}

class(new_cps) <- c(class(new_cps), "cp_refits")




for (x in c("clear", "sef")) {
    opts[[x]]$data = quote(clear_df_bm_)
    opts[[x]]$rcond_threshold <- 1e-3
    opts[[x]]$boot <- NULL
    opts[[x]]$keep_boots <- NULL
}; rm(x)

clear_df_bm_ <- clear_df_bm
x <- "clear"
# How many intially failed to converge?
clear_sef[[x]]$bootstrap$mats %>% length()
# Re-fit ones that failed
clear_refits <- map(clear_sef[[x]]$bootstrap$mats,
       function(X) {
           for (i in 1:ncol(X)) {
               clear_df_bm_[, paste(opts[[x]]$traits)[-1][i]] <- X[,i]
           }
           do.call(cor_phylo, opts[[x]])
       })

# How many still failed?
map_lgl(clear_refits, ~ .x$convcode != 0) %>% 
    sum()

still_failed <- map_lgl(clear_refits, ~ .x$convcode != 0) %>% 
    which()

for (x in c("clear", "sef")) {
    opts[[x]]$rcond_threshold <- 5e-3
}; rm(x)

x <- "sef"
map(still_failed, 
    function(i) {
        X <- clear_sef[[x]]$bootstrap$mats[[i]]
        for (i in 1:ncol(X)) {
            clear_df_bm_[, paste(opts[[x]]$traits)[-1][i]] <- X[,i]
        }
        do.call(cor_phylo, opts[[x]])
    })






```



P-values for including body mass for clearance and SEF, respectively:

```{r clear_sef_pval, echo = FALSE}
cat(sprintf('P = %.3f', pval(clear_sef$clear, "B0")["log_clear_log_mass",]))
cat(sprintf('P = %.3f', pval(clear_sef$sef, "B0")["log_sef_log_mass",]))
```


```{r cor_phylo_write, echo = FALSE, eval = run_models}
write_rds(clear_sef, 'output/inc_mass_cor_phylo.rds')
```



# Session info

This outlines the package versions I used for this script.

```{r session_info, echo = FALSE}
devtools::session_info()
```
