---
title: "Test whether to include log_mass in phylogenetic linear regression"
author: "Lucas Nell"
date: "`r Sys.setenv(TZ='America/Chicago'); format(Sys.Date(), '%d %b %Y')`"
output:
  github_document:
    toc: true
    toc_depth: 2
---

```{r setup, include = FALSE, cache = FALSE, purl = FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_knit$set(root.dir = normalizePath(".."))
```


This file determines whether to use `log_mass` in `phylolm` regressions or not.
These analyses take about half an hour to run.

All p-values below are for whether the coefficient for log(mass) are not equal to zero.


```{r load_packages, echo = FALSE}
suppressPackageStartupMessages({
    library(readr)
    library(dplyr)
    library(tidyr)
    library(purrr)
    library(phylolm)
    library(ape)
})
suppressMessages(devtools::load_all('corphyloCpp'))
invisible(sapply(list.files('R', '*.R', full.names = TRUE), source))
```
```{r load_models, echo = FALSE, eval = TRUE}
diet_fit <- read_rds('output/inc_mass_diet_fit.rds')
absorp_fit <- read_rds('output/inc_mass_absorp_fit.rds')
spp_fits <- read_rds('output/inc_mass_spp_fits.rds')
pos_fits <- read_rds('output/inc_mass_pos_fits.rds')
load('output/inc_mass_corphylo.RData')
```


# Necessary data:

```{r absorp_data}
# Morphometrics by species
spp_df <- get_df('spp')
tr <- get_tr('spp')
# Morphometrics by species and segment
seg_types <- c('prox', 'med', 'dist')
pos_df_list <- lapply(seg_types, get_df, .df = 'pos')
names(pos_df_list) <- seg_types
# Absorption by species
absorp_df <- get_df('absorp')
absorp_se_df <- get_df('absorp', .stat = 'se')  # <-- contains standard errors
absorp_tr <- get_tr('absorp')
# Clearance by species
clear_df <- get_df('clear')
clear_se_df <- get_df('clear', .stat = 'se')  # <-- contains standard errors
clear_tr <- get_tr('clear')
```




# `SEF` on `Diet`

`phylolm` call and output:

```{r sef_diet, eval = FALSE}
set.seed(29851644)
diet_fit <- phylolm(log_sef ~ diet + log_mass, data = spp_df, phy = tr,
                    model = 'lambda', boot = 2000)
```
```{r write_diet, echo = FALSE, eval = FALSE}
write_rds(diet_fit, 'output/inc_mass_diet_fit.rds')
```

P-value:

```{r sef_diet_pval, echo = FALSE}
cat(sprintf('P = %.4g\n', pval(diet_fit, 'log_mass')))
```


# `Absorption` on `Clade`


`phylolm` call and output:

```{r absorp_clade, eval = FALSE}
set.seed(1092141389)
absorp_fit <- suppressWarnings(  # gives warning about lambda being very low
    phylolm(log_absorp ~ clade + log_mass, data = absorp_df, phy = absorp_tr, 
            model = 'lambda', boot = 2000)
)
```
```{r write_absorp, echo = FALSE, eval = FALSE}
write_rds(absorp_fit, 'output/inc_mass_absorp_fit.rds')
```

P-value:

```{r absorp_clade_pval, echo = FALSE}
cat(sprintf('P = %.4g\n', pval(absorp_fit, 'log_mass')))
```



# `Morphometrics` on `Clade`

`phylolm` calls:

```{r sp_analyses, eval = FALSE}
spp_ys <- c("log_intestinal_length", "log_nsa", "log_vill_surface_area",
            "log_total_enterocytes")
set.seed(357885189)
spp_fits <- lapply(
    spp_ys,
    function(y) {
        f <- paste(y, '~ clade + log_mass')
        suppressWarnings(
            do.call("phylolm", list(as.formula(f), data = as.name("spp_df"),
                                    phy = as.name("tr"), model = 'lambda',
                                    boot = 2000))
        )
    })
names(spp_fits) <- spp_ys
```
```{r write_spp, echo = FALSE, eval = FALSE}
write_rds(spp_fits, 'output/inc_mass_spp_fits.rds')
```

P-values:

```{r sp_analyses_pvals, echo = FALSE}
cbind(P = sapply(spp_fits, pval, 'log_mass'))
```





# `Morphometrics` on `Clade`, separately by segment


`phylolm` calls:


```{r pos_analyses, eval = FALSE}
pos_ys <- c('log_intestinal_diameter', 'log_villus_height', 'villus_width', 
            'crypt_width', 'log_sef', 'enterocyte_diameter', 'log_enterocyte_density')
set.seed(632929430)
pos_fits <- lapply(
    seg_types,
    function(pos) {
        # Assigning to obj named <pos>_df so that the call identifies the position
        assign(paste0(pos, '_df'), pos_df_list[[pos]])
        lapply(
            pos_ys,
            function(y) {
                f <- paste(y, ' ~ clade + log_mass')
                arg_list <- list(
                    as.formula(f),
                    data = as.name(paste0(pos, "_df")),
                    phy = as.name("tr"), model = 'lambda')
                # Some models don't find the peak likelihood unless specifying a
                # starting value of 0.1.
                LL_nostart <- suppressWarnings(do.call("phylolm", arg_list))$logLik
                LL_wstart <- suppressWarnings(do.call(
                    "phylolm", c(arg_list, starting.value = 0.1)))$logLik
                if (LL_wstart > LL_nostart) arg_list <- c(arg_list, starting.value = 0.1)
                # Adding number of bootstrap replicates
                arg_list <- c(arg_list, boot = 2000)
                # Now call phylolm
                suppressWarnings(do.call("phylolm", arg_list))
            })
    })
names(pos_fits) <- seg_types
for (i in 1:length(pos_fits)) names(pos_fits[[i]]) <- pos_ys; rm(i)
```
```{r write_pos, echo = FALSE, eval = FALSE}
write_rds(pos_fits, 'output/inc_mass_pos_fits.rds')
```


P-values:

```{r pos_analyses_pvals, echo = FALSE}
cat("proximal:\n")
cbind(P = sapply(pos_fits$prox, pval, .parameters = 'log_mass'))
cat("medial:\n")
cbind(P = sapply(pos_fits$med, pval, .parameters = 'log_mass'))
cat("distal:\n")
cbind(P = sapply(pos_fits$dist, pval, .parameters = 'log_mass'))
```








# `Clearance` and `SEF`

> P-values in all following sections are for whether the coefficient for the `U` 
> matrix is not zero.
> There are two p-values because I'm including the `U` matrix separately
> for the first and second `X` matrix parameters.


```{r clear_sef, eval = FALSE}
# Function to retrieve the U coefficient(s) from a corphylo object
get_U <- function(cp_obj) {
    rn <- rownames(cp_obj$B)[grepl('\\.1', rownames(cp_obj$B))]
    uc <- matrix(as.numeric(cp_obj$B[rn,]), nrow = 1)
    colnames(uc) <- rn
    return(uc)
}

Xmat <- cp_mat(clear_df, c('log_sef', 'log_clear'))
MEmat <- cp_mat(clear_se_df, c('log_sef', 'log_clear'))

# For this comparison, I have to remove one row that doesn't have log_mass
Xmat <- Xmat[!is.na(clear_df$log_mass),]
MEmat <- MEmat[!is.na(clear_df$log_mass),]
Umat <- list( cbind(clear_df$log_mass[!is.na(clear_df$log_mass)]), NULL)
rownames(Umat[[1]]) <- rownames(Xmat)
Umat2 <- list(NULL, cbind(clear_df$log_mass[!is.na(clear_df$log_mass)]))
rownames(Umat2[[2]]) <- rownames(Xmat)

clear_tr <- ape::drop.tip(
    clear_tr,
    tip = clear_tr$tip.label[!clear_tr$tip.label %in% rownames(Xmat)])

# corphylo_cpp run with bootstrapping (takes ~1 min)
set.seed(1844365955)
clear_sef <- corphylo_cpp(X = Xmat, phy = clear_tr, SeM = MEmat, U = Umat, 
                          boot = 2000, n_cores = 4, boot_out = get_U)
clear_sef2 <- corphylo_cpp(X = Xmat, phy = clear_tr, SeM = MEmat, U = Umat2, 
                          boot = 2000, n_cores = 4, boot_out = get_U)
```

P-values:
```{r clear_sef_pval, echo = FALSE}
cat(sprintf('P = %.4g\n', pval(clear_sef)))
cat(sprintf('P = %.4g\n', pval(clear_sef2)))
```


# `Clearance` and `log_enterocyte_density`


```{r clear_ed, eval = FALSE}
Xmat <- cp_mat(clear_df, c('log_enterocyte_density', 'log_clear'))
Xmat <- Xmat[!is.na(rowSums(Xmat)),]

MEmat <- cp_mat(clear_se_df, c('log_enterocyte_density', 'log_clear'))
MEmat <- MEmat[!is.na(rowSums(MEmat)),]

# Fit and bootstrap r (takes ~1 min)
set.seed(1442148819)
clear_ed <- corphylo_cpp(X = Xmat, phy = clear_tr, SeM = MEmat, U = Umat, 
                          boot = 2000, n_cores = 4, boot_out = get_U)
clear_ed2 <- corphylo_cpp(X = Xmat, phy = clear_tr, SeM = MEmat, U = Umat2, 
                          boot = 2000, n_cores = 4, boot_out = get_U)
```

P-values:
```{r clear_ed_pval, echo = FALSE}
cat(sprintf('P = %.4g\n', pval(clear_ed)))
cat(sprintf('P = %.4g\n', pval(clear_ed2)))
```


# `Absorption` and `log_total_enterocytes`


```{r absorp_te, eval = FALSE}
Xmat <- cp_mat(absorp_df, c('log_absorp', 'log_total_enterocytes'))
MEmat <- cp_mat(absorp_se_df, c('log_absorp', 'log_total_enterocytes'))

Umat <- list(cbind(absorp_df$log_mass[!is.na(absorp_df$log_mass)]), NULL)
rownames(Umat[[1]]) <- rownames(Xmat)
Umat2 <- list(NULL, cbind(absorp_df$log_mass[!is.na(absorp_df$log_mass)]))
rownames(Umat2[[2]]) <- rownames(Xmat)

# Fit and bootstrap
set.seed(2016097648)
absorp_te <- corphylo_cpp(X = Xmat, phy = absorp_tr, SeM = MEmat, U = Umat, 
                          boot = 2000, n_cores = 4, boot_out = get_U)
# # This one gives numerical issues: non positive definite correlation matrix
# absorp_te2 <- corphylo_cpp(X = Xmat, phy = absorp_tr, SeM = MEmat, U = Umat2, 
#                           boot = 2000, n_cores = 4, boot_out = get_U)
```

P-value:
```{r absorp_te_pval, echo = FALSE}
cat(sprintf('P = %.4g\n', pval(absorp_te)))  #; pval(absorp_te2)
```


```{r corphylo_write, eval = FALSE}
save(clear_sef, clear_sef2, clear_ed, clear_ed2, absorp_te, 
     file = 'output/inc_mass_corphylo.RData')
```



# Session info

This outlines the package versions I used for this script.

```{r session_info, echo = FALSE}
devtools::session_info()
```
