

<!---
% This document talks about and produces a phylogenetic tree for the supplemental info.
-->

The phylogenetic tree was downloaded from [`timetree.org`](timetree.org)
\citep{Kumar_2017}.
Not every species in our phylogeny was used for every analysis.
Analyses related to morphometric, clearance, and absorption data had separate
datasets consisting of different species.
Figure \ref{fig:phylo-plot} shows the full phylogeny and which species were 
present in each dataset.


```{r phylo-packages-data, echo = FALSE}
suppressPackageStartupMessages({
    library(ape)
    library(dplyr)
    library(purrr)
    library(ggplot2)
    library(ggtree)
    library(ggimage)
    library(grid)
})
invisible(sapply(list.files('R', '*.R', full.names = TRUE), source))
tr <- get_tr('spp') # <-- full tree
morph_df <- get_df('spp')
absorp_df <- get_df('absorp')
clear_df <- get_df('clear')

# Data frame of which datasets each species is in:
data_df <- morph_df %>%
    distinct(species) %>% 
    mutate(`morph.` = TRUE, 
           `absorp.` = species %in% absorp_df$species,
           `clear.` = species %in% clear_df$species)
rownames(data_df) <- data_df$species
data_df <- data_df[, -1]
```




```{r phylo-plot, echo = FALSE, fig.cap = "Phylogeny of species used in these analyses. Boxes on the right indicate whether the species was used in morphometric (\"morph.\"), absorption (\"absorp.\"), or clearance (\"clear.\") analyses. Gray boxes indicate a species was absent. The scale is in millions of years."}
gg_tr <- ggtree(tr)
gg_tr$data$x <- gg_tr$data$x - max(gg_tr$data$x)
gg_tr <- gg_tr + 
    theme_tree(axis.title.x = element_blank(),
               plot.margin = margin(0,0,0,0)) +
    geom_tiplab(aes(x = x + 2), size = 3, fontface = 'bold.italic')

suppressMessages(
    gg_tr <- gheatmap(gg_tr, data_df, offset = 38, width = 0.2, font.size = 2.5, 
                      colnames_position ='bottom', colnames_angle = -30, 
                      colnames_offset_y = -0.25, colnames_offset_x = 2) +
        scale_fill_manual(breaks = c(TRUE, FALSE),  values = c("gray70", "dodgerblue"),
                          guide = FALSE) +
        geom_treescale(x = -90, y = 15, offset = 0.5, width = 10)
)

img_ids <- c("18bfd2fc-f184-4c3a-b511-796aafcc70f6",  # rodent
             "13ddd312-92ad-4873-9d65-4732b86cfbbd")  # bat
imgs <- map(img_ids, ~ rphylopic::image_data(.x, 512)[[1]])
img_credits <- map_chr(img_ids, ~ rphylopic::image_get(.x, options = "credit")$credit)

gg_tr <- gg_tr +
    rphylopic::add_phylopic(imgs[[1]], color = "black", alpha = 0.25,
                            x = -73, y = 6, ysize = 20) +
    rphylopic::add_phylopic(imgs[[2]], color = "black", alpha = 0.25,
                            x = -60, y = 13, ysize = 20) +
    # Credit at the bottom
    labs(caption = sprintf("Silhouettes: %s, %s", img_credits[1], img_credits[2]))

gg_tr
```

