

Table \ref{tab:results-clear-sef-kable}
refers to `corphylo` estimates for the correlation between clearance and SEF.
The table includes the Pearson correlation coefficient ($r$) between the two traits 
and phylogenetic signals from an Ornstein-Uhlenbeck process for each individual trait.


```{r results-clear-sef-kable}
filter(summ_df, model == 'corphylo')$data[[1]] %>% 
    mutate(Y = str_replace_all(Y, 'log_', '') %>% str_replace_all('_', ' '), 
           parameter = c('$r$', sprintf('$d_{\\text{%s}}$', gsub('log_', '', Y[2])), 
                         sprintf('$d_{\\text{%s}}$', gsub('log_', '', Y[3])))) %>% 
    select(-phy_model, -Y) %>% 
    pretty_df(., caption = str_c('\\texttt{corphylo} model summary for the ',
                                 'correlation between clearance and SEF. ',
                                 '$r$ indicates the correlation and $d_X$ indicates ',
                                 'the phylogenetic signal from an Ornstein-Uhlenbeck ',
                                 'process for variable $X$. ',
                                 'Both variables were log-transformed for this ',
                                 'analysis.'))
```




### Influential value for correlation between clearance and SEF

#### Jackknifing

We used jackknifing to determine how influential each point is to the correlation 
estimate in the `corphylo` model for `clearance ~ SEF`.
See equation \ref{eq:jackknife} for how I calculated influence absolute Z-scores.
Figure \ref{fig:results-influence-plot-cp} shows that one species 
(*Tadarida brasiliensis*) has a disproportionate effect on the correlation.


```{r results-influence-jack-corphylo}
# (See `R/model_summaries.R` for the `jack_corphylo` function.)
clear_df <- get_df('clear')
clear_se_df <- get_df('clear', .stat = 'se')
clear_tr <- get_tr('clear')
cp_jack <- jack_corphylo(cp_mod, mean_df = clear_df, se_df = clear_se_df,
                         phy = clear_tr, par_names = c('log_sef', 'log_clear'))
```



```{r results-influence-plot-cp, fig.height = 3, fig.width = 6.5, fig.cap = "Influential points for the correlation (Pearson's $r$) between clearance and SEF. (A) Clearance against SEF where point sizes are proportional to their influence absolute Z-score, color indicates clade, and both axes are on the log scale. (B) Influence absolute Z-scores by species."}
# Adding morphometric data to the jackknife data frame
cor_jack_df <- dplyr::left_join(cp_jack, get_df('clear'), by = 'species') %>% 
    select(diet, clade, species, log_sef, log_clear, influence) %>% 
    mutate(species = reorder(species, abs(influence)))

cor_jack_p <- list()

# Plot showing context of influential value
cor_jack_p[['A']] <- cor_jack_df %>%
    # Exponentiate variables for plotting so we can more transparently
    # provide axis labels of the non-transformed numbers.
    # The same is done below when plotting the raw data.
    mutate_at(vars(log_sef, log_clear), exp) %>%
    ggplot(aes(log_sef, log_clear, color = clade, size = abs(influence))) +
    geom_point() +
    scale_x_continuous("Surface enlagement factor (SEF)",
                       trans = 'log', breaks = 8 * 1.5^(0:2)) +
    scale_y_continuous(bquote(.("L-arabinose clearance (µ") *
                                  l ~ min^{-1} ~ cm^{-2} * ")"),
                       trans = 'log', breaks = 1 * 3^(0:2)) +
    scale_size_area('Influence\n|Z-score|:', breaks = 1:2, max_size = 4) +
    scale_color_manual('Clade:', values = c('#9ecae1', '#de2d26')) +
    theme(legend.position = 'left',
          legend.title = element_text(size = 10, face = 'bold.italic'), 
          legend.margin = margin(0,0,0,0))


# Plotting to show better the actual |Z-scores|
cor_jack_p[['B']] <- cor_jack_df %>%
    ggplot(aes(abs(influence), species)) +
    geom_segment(aes(yend = species), xend = 0, color = 'gray40') +
    geom_point(size = 2, color = 'dodgerblue') +
    theme(axis.title.y = element_blank(),
          axis.text.y = element_text(face = 'italic'),
          plot.margin = margin(5.5, 5.5, 5.5, 8.5)) +
    xlab("Influence |Z-score|")

cowplot::plot_grid(plotlist = cor_jack_p, ncol = 2, rel_widths = c(5, 4),
                   labels = LETTERS[1:2], label_fontface = 'plain')
```





#### Re-running model without influential point

```{r results-influence-rerun-corphylo, cache = TRUE}
Xmat <- cp_mat(clear_df, c('log_sef', 'log_clear'))
Xmat <- Xmat[clear_df$species != 'Tadarida brasiliensis',]
MEmat <- cp_mat(clear_se_df, c('log_sef', 'log_clear'))
MEmat <- MEmat[clear_se_df$species != 'Tadarida brasiliensis',]
tr_ <- clear_tr %>% filter_tr(rownames(Xmat))
set.seed(1052373084)
influence_rerun <- corphylo_cpp(X = Xmat, phy = tr_, SeM = MEmat,
                                boot = 2000, n_cores = 4)
```


Table \ref{tab:results-re-run-summ-dfs-corphylo} shows the `corphylo` output
when this influential point is excluded.

```{r results-re-run-summ-dfs-corphylo}
rerun_df <- data_frame(parameter = c('$r$', '$d_{\\text{sef}}$', 
                                     '$d_{\\text{clear}}$'),
                       estimate = c(influence_rerun$cor.matrix[1,2], 
                                    influence_rerun$d),
                       lower = apply(influence_rerun$bootstrap, 2, 
                                     quantile, probs = 0.025),
                       upper = apply(influence_rerun$bootstrap, 2, 
                                     quantile, probs = 0.975),
                       P = c(pval(influence_rerun)[1], rep("–", 2)))
rerun_df %>% 
    pretty_df(., caption = str_c('\\texttt{corphylo} model summary for the ',
                                 'correlation between clearance and SEF ',
                                 'without the most influential point. ',
                                 '$r$ indicates the correlation and $d_X$ indicates ',
                                 'the phylogenetic signal from an Ornstein-Uhlenbeck ',
                                 'process for variable $X$. ',
                                 'Both variables were log-transformed for this ',
                                 'analysis.'))
```




