

Table \ref{tab:results-clear-sef-kable}
refers to `cor_phylo` estimates for the correlation between clearance and SEF.
The table includes the Pearson correlation coefficient ($r$) between the two traits 
and phylogenetic signals from an Ornstein-Uhlenbeck process for each individual trait.


```{r results-clear-sef-kable}
filter(summary_df, model == 'cor_phylo')$data[[1]] %>% 
    mutate(Y = str_replace_all(Y, 'log_', '') %>% str_replace_all('_', ' '), 
           parameter = c('$r$', sprintf('$d_{\\text{%s}}$', gsub('log_', '', Y[2])), 
                         sprintf('$d_{\\text{%s}}$', gsub('log_', '', Y[3])))) %>% 
    select(-phy_model, -Y) %>% 
    pretty_df(., caption = str_c('\\texttt{cor\\_phylo} model summary for the ',
                                 'correlation between clearance and SEF. ',
                                 '$r$ indicates the correlation and $d_X$ indicates ',
                                 'the phylogenetic signal from an Ornstein-Uhlenbeck ',
                                 'process for variable $X$. ',
                                 'Both variables were log-transformed for this ',
                                 'analysis.'))
```




### Influential value for correlation between clearance and SEF

#### Jackknifing

We used jackknifing to determine how influential each point is to the correlation 
estimate in the `cor_phylo` model for `clearance ~ SEF`.
See equation \ref{eq:jackknife} for how I calculated influence absolute Z-scores.
Figure \ref{fig:results-influence-plot-cp} shows that one species 
(*Tadarida brasiliensis*) has a disproportionate effect on the correlation.


```{r results-influence-jack-cor-phylo}
# (See `R/model_summaries.R` for the `jack` function.)
# Making data frame with both means and SEs:
clear_sef_df <- bind_cols(get_df('clear') %>% select(species, log_sef, log_clear),
                          get_df('clear', .stat = 'se') %>% select(log_sef, log_clear) %>% 
                              rename(log_sef_se = log_sef, log_clear_se = log_clear))
clear_tr <- get_tr('clear')
cp_jack <- jack(cp_mod)
```



```{r results-influence-plot-cp, fig.height = 3, fig.width = 6.5, fig.cap = "Influential points for the correlation (Pearson's $r$) between clearance and SEF. (A) Clearance against SEF where point sizes are proportional to their influence absolute Z-score, color indicates clade, and both axes are on the log scale. (B) Influence absolute Z-scores by species."}
# Adding morphometric data to the jackknife data frame
cor_jack_df <- dplyr::left_join(cp_jack, get_df('clear'), by = 'species') %>% 
    select(diet, clade, species, log_sef, log_clear, influence) %>% 
    mutate(species = reorder(species, abs(influence)))

cor_jack_p <- list()

# Plot showing context of influential value
cor_jack_p[['A']] <- cor_jack_df %>%
    # Exponentiate variables for plotting so we can more transparently
    # provide axis labels of the non-transformed numbers.
    # The same is done below when plotting the raw data.
    mutate_at(vars(log_sef, log_clear), exp) %>%
    ggplot(aes(log_sef, log_clear, color = clade, size = abs(influence))) +
    geom_point() +
    scale_x_continuous("Surface enlagement factor (SEF)",
                       trans = 'log', breaks = 8 * 1.5^(0:2)) +
    scale_y_continuous(bquote(.("L-arabinose clearance (µ") *
                                  l ~ min^{-1} ~ cm^{-2} * ")"),
                       trans = 'log', breaks = 1 * 3^(0:2)) +
    scale_size_area('Influence\n|Z-score|:', breaks = 1:2, max_size = 4) +
    scale_color_manual('Clade:', values = c('#9ecae1', '#de2d26')) +
    theme(legend.position = 'left',
          legend.title = element_text(size = 10, face = 'bold.italic'), 
          legend.margin = margin(0,0,0,0))


# Plotting to show better the actual |Z-scores|
cor_jack_p[['B']] <- cor_jack_df %>%
    ggplot(aes(abs(influence), species)) +
    geom_segment(aes(yend = species), xend = 0, color = 'gray40') +
    geom_point(size = 2, color = 'dodgerblue') +
    theme(axis.title.y = element_blank(),
          axis.text.y = element_text(face = 'italic'),
          plot.margin = margin(5.5, 5.5, 5.5, 8.5)) +
    xlab("Influence |Z-score|")

cowplot::plot_grid(plotlist = cor_jack_p, ncol = 2, rel_widths = c(5, 4),
                   labels = LETTERS[1:2], label_fontface = 'plain')
```





#### Re-running model without influential point

```{r results-influence-rerun-cor-phylo, cache = TRUE}
clear_sef_df_ <- clear_sef_df %>% 
    filter(species != 'Tadarida brasiliensis')
clear_tr_ <- clear_tr %>% filter_tr(clear_sef_df_$species)
call_ <- cp_mod$call
call_$data <- quote(clear_sef_df_)
call_$phy <- quote(clear_tr_)
# This helps bootstrap estimates to be more numerically stable:
call_$rcond_threshold <- 1e-4

set.seed(1052373084)
influence_rerun <- eval(call_)
```


```{r cor_phylo-influence-rerun-refits}
# Some bootstrap replicates did not converge, so I'm refitting them using a higher
# threshold for the reciprocal condition number of two matrices inside the
# likelihood function (the `rcond_threshold` argument to `cor_phylo`).
# This makes the optimization process more strongly "bounce away" from badly
# conditioned matrices.
# From trial and error, using a `rcond_threshold` value of `2e-3` seemed to make
# all the non-converging replicates converge and provide sensible results.

cp_rerun_boot_refits <- refit_boots(influence_rerun, rcond_threshold = 2e-3)
```



Table \ref{tab:results-re-run-summ-dfs-cor-phylo} shows the `cor_phylo` output
when this influential point is excluded.

```{r results-re-run-summ-dfs-cor-phylo}
bci <- boot_ci(influence_rerun, cp_rerun_boot_refits)
rerun_df <- data_frame(parameter = c('$r$', '$d_{\\text{sef}}$', 
                                     '$d_{\\text{clear}}$'),
                       estimate = c(influence_rerun$corrs[1,2], 
                                    influence_rerun$d),
                       lower = c(bci$corrs[2,1],
                                 bci$d[,'lower']),
                       upper = c(bci$corrs[1,2],
                                 bci$d[,'upper']),
                       P = c(pval(influence_rerun, "corrs", cp_rerun_boot_refits)[1,2], 
                             rep("–", 2)))
rerun_df %>% 
    pretty_df(., caption = str_c('\\texttt{cor\\_phylo} model summary for the ',
                                 'correlation between clearance and SEF ',
                                 'without the most influential point. ',
                                 '$r$ indicates the correlation and $d_X$ indicates ',
                                 'the phylogenetic signal from an Ornstein-Uhlenbeck ',
                                 'process for variable $X$. ',
                                 'Both variables were log-transformed for this ',
                                 'analysis.'))
```




