---
title: 'Electronic Supplementary Material: Supplemental Methods and Results'
subtitle: "``Morphometric bases for intestinal paracellular absorption in bats and rodents''"
date: "`r Sys.setenv(TZ='America/Chicago'); format(Sys.Date(), '%d %b %Y')`"
fontsize: 12pt
csl: "ecology.csl"
bibliography: "references.bib"
geometry: margin=1in,letterpaper
documentclass: article
tables: true
graphics: true
colorlinks: true
output:
    bookdown::pdf_document2:
        fig_caption: yes
        number_sections: yes
        template: template.tex
        toc: yes
        toc_depth: 2
---

```{r setup, include=FALSE, cache = FALSE}
suppressPackageStartupMessages({
    library(knitr)
    library(knitcitations)
    library(bookdown)
})
opts_chunk$set(echo = FALSE, cache = FALSE, dev = 'quartz_pdf', fig.pos = 'h')
opts_knit$set(root.dir = normalizePath(".."))
knit_theme$set(knit_theme$get("acid"))
cleanbib()
options("citation_format" = "pandoc")
```
 
This section displays all results from analyses using
`phylolm::phylolm` `r citep(citation("phylolm"))` and 
a modified version of `ape::corphylo`
`r citep(list(citation("ape"), "10.1111/j.1365-2435.2009.01596.x"))`
that provides parametric bootstrapping.

In the tables below, the "parameter" column indicates which variable (either an X 
variable or phylogenetic signal parameter) that row's numbers refer to.
Column "estimate" is the maximum likelihood estimate for the specified parameter, 
while columns "lower" and "upper" are lower and upper bounds of the 95% confidence 
interval for the coefficient estimate.
"P" is the p-value for an estimate not being equal to zero.
Both p-values and CIs were obtained by parametric bootstrapping of the `phylolm` model.
We do not present P-values for phylogenetic signals because they were bound
above zero by the models; thus confidence intervals are more informative.

Within a given table, multiple parameter-column values for a given Y do not indicate 
that separate analyses were performed for each parameter, merely that both 
were included in the same model.


 
```{r libraries, echo = FALSE, cache = FALSE}
# Packages needed for this script 
suppressPackageStartupMessages({
    library(readr)
    library(dplyr)
    library(tidyr)
    library(stringr)
    library(purrr)
    library(phylolm)
    library(ape)
    library(ggplot2)
    library(grid)
    library(gridExtra)
})
# Set default ggplot2 theme
theme_set(theme_classic())
# Custom package that does bootstrapping on corphylo object
suppressMessages(devtools::load_all('corphyloCpp'))
# Sourcing scripts from R directory
invisible(sapply(list.files('R', '*.R', full.names = TRUE), source))

# Function to print data frames for LaTeX output using booktabs package
pretty_df <- function(df_, caption = NULL) {
    kable_obj <- knitr::kable(df_, format = 'latex', booktabs = TRUE, caption = caption)
    kable_str <- unlist(str_split(kable_obj, '\n'))
    kable_str <- kable_str[kable_str != "\\addlinespace"]
    for (i in 1:(length(kable_str)-1)) {
        kable_str[i] <- gsub("textbackslash\\{\\}", "", kable_str[i])
        kable_str[i] <- gsub("\\\\textasciicircum\\{\\}", "\\^", kable_str[i])
        kable_str[i] <- gsub("\\\\\\$", "$", kable_str[i])
        kable_str[i] <- gsub("\\\\\\{", "{", kable_str[i])
        kable_str[i] <- gsub("\\\\\\}", "}", kable_str[i])
        kable_str[i] <- gsub("beta\\\\_", "beta_", kable_str[i])
        # kable_str[i] <- gsub("\\\\_", "beta_", kable_str[i])
        if (grepl('^ &', kable_str[i]) & !grepl('^ &', kable_str[(i+1)]) & 
            !grepl('^\\\\bottomrule', kable_str[(i+1)])) {
            kable_str[i] <- str_c(kable_str[i], "\n", "\\addlinespace")
        }
    }
    new_kable_str <- str_c(kable_str, collapse = '\n')
    kable_obj[1] <- new_kable_str
    return(kable_obj)
}
``` 
 
 
```{r parse-output, echo = FALSE} 
summ_df <- read_csv('output/models_summaries.csv', col_types = 'ccccdddd') %>%
    mutate(X = ifelse(X == 'cladeBat', 'clade', X),
           pos = ifelse(is.na(pos), 'all', pos)) %>%  
    rename(estimate = value) %>%
    select(pos, Y, X, estimate, lower, upper, P) %>% 
    mutate_if(is.numeric, function(s) sprintf('%.4g', s)) %>% 
    mutate(model = ifelse((X %in% c('absorp', 'log_clear', 'd') | 
                               Y %in% c('absorp', 'log_clear')) & 
                              !X %in% c('clade', 'lambda', 'log_mass'), 
                          'corphylo','phylolm'),
           P = ifelse(X %in% c('lambda', 'd', 'alpha'), "–", P),
           # Explicitly ordering parameters:
           X = factor(X, 
                      levels = c("clade", "log_mass", 
                                 "dietOmnivorous", "dietProtein", 
                                 "log_clear", "log_total_enterocytes", 
                                 "lambda", "alpha", "d"))) %>% 
    rename(parameter = X) %>% 
    nest(-model, -pos)
plm_mods <- list(absorp = read_rds('output/models_absorp.rds'),
               diet = read_rds('output/models_diet.rds'),
               pos = read_rds('output/models_pos.rds'),
               spp = read_rds('output/models_spp.rds'))
cp_mods <- read_rds('output/models_corphylo.rds')
``` 
 

## `phylolm` estimates

Tables \ref{tab:spp-summs}--\ref{tab:dist-summs} summarize the coefficient estimates
for all `phylolm` model fits.

The list below describes all possible "parameter" column values:

- log_absorp: fractional absorption
- log_intestinal_length: intestinal length
- log_nsa: nominal surface area
- log_sef: surface enlargement factor
- log_total_enterocytes: total enterocytes
- log_vill_surface_area: total villous surface area
- clade: clade (bat or rodent)
- log_mass: body mass
- dietOmnivorous: binary variable whose value is 1 if the species is omnivorous
- dietProtein: binary variable whose value is 1 if the species is carnivorous
- lambda: phylogenetic correlation parameter (Pagel's lambda)


```{r spp-summs} 
filter(summ_df, pos == 'all', 
                          model == 'phylolm')$data[[1]] %>% 
    arrange(Y, parameter) %>%
    mutate(Y = ifelse(Y == lag(Y, default = 'XXX'), "", Y)) %>% 
    pretty_df(., caption = '\\texttt{phylolm} model output for analyses by species only.')
``` 


```{r prox-summs} 
filter(summ_df, pos == 'prox', model == 'phylolm')$data[[1]] %>% 
    arrange(Y, parameter) %>%
    mutate(Y = ifelse(Y == lag(Y, default = 'XXX'), "", Y)) %>% 
    pretty_df(., caption = str_c('\\texttt{phylolm} model output for proximal intestinal ',
                                 'segment analyses.'))
``` 


```{r med-summs} 
filter(summ_df, pos == 'med', model == 'phylolm')$data[[1]] %>% 
    arrange(Y, parameter) %>%
    mutate(Y = ifelse(Y == lag(Y, default = 'XXX'), "", Y)) %>% 
    pretty_df(., caption = str_c('\\texttt{phylolm} model output for medial intestinal ',
                                 'segment analyses.'))
``` 


```{r dist-summs}
filter(summ_df, pos == 'dist', model == 'phylolm')$data[[1]] %>% 
    arrange(Y, parameter) %>%
    mutate(Y = ifelse(Y == lag(Y, default = 'XXX'), "", Y)) %>% 
    pretty_df(., caption = str_c('\\texttt{phylolm} model output for distal intestinal ',
                                 'segment analyses.'))
``` 




## `corphylo` estimates

Except when the parameter column is "d", rows below refer to the correlation between the
"Y" and "parameter" variables. There are no real X and Y variables here.

When the parameter is "d", this row refers to the phylogenetic signal from an 
Ornstein-Uhlenbeck process for variable "Y".

### `Clearance` and `SEF`

```{r clear-sef-kable}
filter(summ_df, model == 'corphylo')$data[[1]][1:3,] %>% 
    pretty_df(.)
```

### `Clearance` and `log_enterocyte_density`

```{r clear-ed-kable}
filter(summ_df, model == 'corphylo')$data[[1]][4:6,] %>% 
    pretty_df(.)
```


### `Absorption` and `log_total_enterocytes`

```{r absorp-te-kable}
filter(summ_df, model == 'corphylo')$data[[1]][7:9,] %>% 
    pretty_df(.)
```





## Influential values


### Jackknifing

In the `phylolm` model for `log_intestinal_length ~ clade`, one point is very
influential to the results, which makes me less confident in the conclusions from the
full dataset.
Below we use jackknifing to determine how influential each point is to the `log_mass`
and `clade` estimates.

```{r influence-jack, cache = TRUE}
# (See `R/model_summaries.R` for the `jack_phylolm` function.)
int_jack <- jack_phylolm(plm_mods$spp$log_intestinal_length, 
                         phy = get_tr('spp'))
```


Figure \ref{fig:influence-plots} shows that one species (*Artibeus lituratus*) has
a disproportionate effect on all three coefficient estimates.


```{r influence-plots, fig.height = 7, fig.width = 4, fig.cap = "Plot of influence values for regression of intestinal length on clade and body mass. Point sizes are proportional to their influence (normalized to range from 0 to 1) on the maximum likelihood estimate for (A) clade, (B) body mass, and (C) intercept. Color indicates clade."}
# Keep the same area options for every plot
area_opts <- scale_size_area(breaks = c(0.5, 1.0), max_size = 4)

get_infl <- function(species, .est) {
    max_infl <- max(abs({int_jack %>% filter(estimate == .est)}$influence))
    sapply(species, 
           function(s) {
               infl <- abs({int_jack %>% 
                       filter(species == s, estimate == .est)}$influence[1])
               return(infl / max_infl)
           })
}

add_title <- function(.p, .title) {
    if (.p$labels$x == 'pos') {
        suppressMessages({
            .p <- .p + 
                scale_x_continuous(breaks = 1:3, limits = c(0.4, 3.43),
                                   labels = c('Proximal', 'Medial', 'Distal'))
        })
    }
    x_range <- ggplot_build(.p)$layout$panel_ranges[[1]]$x.range
    y_range <- ggplot_build(.p)$layout$panel_ranges[[1]]$y.range
    if (!is.null(.p$coordinates$trans$x)) {
        x_range <- .p$coordinates$trans$x$inverse(x_range)
    }
    if (!is.null(.p$coordinates$trans$y)) {
        y_range <- .p$coordinates$trans$y$inverse(y_range)
    }
    min_x <- min(x_range) + 0.02 * diff(x_range)
    max_y <- max(y_range) - 0.02 * diff(y_range)
    .p <- .p + 
        geom_text(data = NULL, label = .title, 
                  x = min_x, y = max_y, hjust = 0, vjust = 1, 
                  size = 14 * (25.4/72), fontface = 'plain', color = 'black')
    return(.p)
}


infl_plots <- function(estimate, guide_area = TRUE, guide_color = TRUE, title_ = NULL) {

    p <- get_df('spp') %>%
        # Exponentiate variables for plotting so we can more transparently
        # provide axis labels of the non-transformed numbers.
        # The same is done below when plotting the raw data.
        mutate_at(vars(log_mass, log_intestinal_length), exp) %>%
        mutate(influence = get_infl(species, estimate)) %>% 
        ggplot(aes(log_mass, log_intestinal_length, color = clade, size = influence)) +
        geom_point() +
        # All these plots include log(mass) on the x-axis so we're including this here.
        scale_x_continuous('Body mass (g)', trans = 'log', breaks = 10 * 4^(0:2),
                           limits = exp(c(2.043877, 5.199205)), expand = c(0,0)) +
        scale_y_continuous('Intestinal length (cm)', trans = 'log', breaks = 10*2^(0:2)) +
        # xlab('log(body mass [g])') +
        # ylab('log(intestinal length [cm])') +
        theme(plot.title = element_text(size = 12))
    if (guide_area) {
        p <- p + scale_size_area(breaks = c(0.5, 1.0), max_size = 4)
    } else {
        p <- p + scale_size_area(breaks = c(0.5, 1.0), max_size = 4, guide = FALSE)
    }
    if (guide_color) {
        p <- p + scale_color_manual(values = c('#9ecae1', '#de2d26'))
    } else {
        p <- p + scale_color_manual(values = c('#9ecae1', '#de2d26'), guide = FALSE)
    }
    if (!is.null(title_)) {
        p <- add_title(p, title_)
    }
    return(p)
}

# Plotting influence on `cladeBat` onto plot of log_intestinal_length ~ log_mass
infl_clade <- infl_plots('cladeBat', guide_area = FALSE, guide_color = FALSE, 'A') +
    ggtitle('clade') +
    theme(axis.title = element_blank())

# Plotting influence on `log_mass` onto plot of log_intestinal_length ~ log_mass
infl_mass <- infl_plots('log_mass', TRUE, TRUE, 'B') +
    ggtitle('body mass') +
    theme(axis.title = element_blank())

# Plotting influence on `(Intercept)` onto plot of log_intestinal_length ~ log_mass
infl_int <- infl_plots('intercept', FALSE, FALSE, 'C') +
    ggtitle('intercept') +
    theme(axis.title = element_blank())

# To extract the legend from a ggplot object
g_legend <- function(a.gplot){
    tmp <- ggplot_gtable(ggplot_build(a.gplot))
    leg <- which(sapply(tmp$grobs, function(x) x$name) == "guide-box")
    legend <- tmp$grobs[[leg]]
    legend
}

.legend <- g_legend(infl_mass)

.ylab <- ggplot(data_frame(x = 0, y = 0), aes(x, y)) +
    geom_text(label = 'Intestinal length (cm)', angle = 90) +
    theme_void()
.xlab <- ggplot(data_frame(x = 0, y = 0), aes(x, y)) +
    geom_text(label = 'Body mass (g)') +
    theme_void()
.void <- ggplot() + geom_blank()


grobs <- grid.arrange(.void, ggplotGrob(infl_clade), .void,
                      .ylab, ggplotGrob(infl_mass + theme(legend.position = 'none')), .legend,
                      .void, ggplotGrob(infl_int), .void,
                      .void, .xlab, .void,
                      ncol = 3, nrow = 4, widths = c(1/10, 7/10, 2/10),
                      heights = c(1, 1, 1, 1/10))

grid.draw(grobs)
```




### Re-running model without influential point

Below is a summary of the model estimates when including all values.

```{r influence-firstrun}
influence_og <- plm_mods$spp$log_intestinal_length
first_run_df <- cbind(cbind(estimate = c(influence_og$coefficients, 
                                         lambda = influence_og$optpar)),
                      t(influence_og$bootconfint95)[c("(Intercept)", "cladeBat", 
                                                      "log_mass", "lambda"), ]) %>% 
    as_tibble %>% 
    rename(lower = `2.5%`, upper = `97.5%`) %>% 
    mutate(parameter = c("(Intercept)", "cladeBat", "log_mass", 
                         "lambda")) %>% 
    select(parameter, everything()) %>% 
    mutate_if(is.numeric, function(x) sprintf('%.3g', x))
```


Now the results without including the most influential value:

```{r influence-rerun, cache = TRUE}
spp_df <- get_df('spp')
spp_tr <- get_tr('spp')
.df <- get_df('spp')
.df <- .df[.df$species != 'Artibeus lituratus',]
.tr <- get_tr('spp') %>% filter_tr(.df$species)
set.seed(1018666821)
suppressWarnings(  # gives warning about low lambda
    influence_rerun <- phylolm(log_intestinal_length ~ clade + log_mass, 
                               model = 'lambda', data = .df, boot = 2000, 
                               phy = .tr))
```

```{r re-run-summ-dfs}
rerun_df <- cbind(cbind(estimate = c(influence_rerun$coefficients, 
                                     lambda = influence_rerun$optpar)),
                  t(influence_rerun$bootconfint95)[c("(Intercept)", "cladeBat", 
                                                     "log_mass", "lambda"), ]) %>% 
    as_tibble %>% 
    rename(lower = `2.5%`, upper = `97.5%`) %>% 
    mutate(parameter = c("(Intercept)", "cladeBat", "log_mass", 
                         "lambda")) %>% 
    select(parameter, everything()) %>% 
    mutate_if(is.numeric, function(x) sprintf('%.3g', x))
list(first_run_df, rerun_df) %>% 
    lapply(function(x) {
        mutate(x, parameter = sapply(
            parameter, 
            function(.p){
                switch(.p, 
                       `(Intercept)` = "$\\beta_0$",
                       cladeBat = "$\\beta_{clade}$",
                       log_mass = "$\\beta_{mass}$",
                       lambda = "$\\lambda$",
                       .p)}))
    }) %>% 
    pretty_df(., caption = str_c('\\texttt{phylolm} model estimates for regression of ',
                                 'log(intestinal length) on clade and log(body mass), ',
                                 'with (left) and  without (right) influential point ',
                                 'present in dataset.'))
```


We can see that we're much less sure about mass's influence on intestinal length when
we remove that one point.
Fortunately it does not change our conclusions regarding the influence of clade
(P-value for $\beta_{\text{clade}}$ with influential point:
`r sprintf("%.4g", pval(influence_og, 'cladeBat'))`;
P without it: `r sprintf("%.4g", pval(influence_rerun, 'cladeBat'))`).

## Percent differences

This is for the percent differences in the discussion and abstract.

We calculated rodent and bat estimates at `log(body mass) = 0` to remove the effect
of body mass.
If the variable was log-transformed (all those below were), then we exponentiated
both rodent and bat estimates before computing the proportional, then percentage, 
difference:

\begin{equation}
  100 \times \frac{ e^{\beta_0 + \beta_{\text{clade}}} - e^{\beta_0} }{ e^{\beta_0} }
\end{equation}


where $\beta_0$ is the intercept and $\beta_{\text{clade}}$ the coefficient for clade.


```{r percent-diff}
perc_diff <- function(.mod) {
    
    # Detect log-transform from formula
    logged <- grepl('^log', paste(.mod$formula)[2])
    
    coefs <- {.mod %>% summary %>% coef}[,'Estimate']
    rodents <- coefs[['(Intercept)']]
    bats <- rodents + coefs[['cladeBat']]
    if (logged) {
        rodents <- exp(rodents)
        bats <- exp(bats)
    }
    prop_diff <- (bats - rodents) / rodents
    return(prop_diff * 100)
}
```


__NSA\:__
```{r percent-diff-nsa}
cat(sprintf('%.4g %%\n', perc_diff(plm_mods$spp$log_nsa)))
```


__SEF\:__
```{r percent-diff-sef}
cat(sprintf('%.4g %%\n', mean(c(
    perc_diff(plm_mods$pos$prox$log_sef),
    perc_diff(plm_mods$pos$med$log_sef),
    perc_diff(plm_mods$pos$dist$log_sef)
))))
```


__Enterocyte density\:__
```{r percent-diff-ent-dens}
cat(sprintf('%.4g %%\n', mean(c(
    perc_diff(plm_mods$pos$prox$log_enterocyte_density), 
    perc_diff(plm_mods$pos$med$log_enterocyte_density), 
    perc_diff(plm_mods$pos$dist$log_enterocyte_density)))))
```




## SEF on Clade and log(Mass)

This is to determine whether there's an effect of body mass on SEF. It appears there
is not.

```{r sef-clade-mass, echo = TRUE, cache = TRUE}
# Retrieve necessary data:
tr <- get_tr('spp')
spp_df <- get_df('spp')
# Run new phylolm fit:
set.seed(940318092)
mod <- suppressWarnings(
    phylolm(log_sef ~ clade + log_mass, data = spp_df, phy = tr, 
            model = "lambda", boot = 2000))
# P-values for coefficient estimates != 0, using parametric 
# bootstrapping:
pval(mod, c('log_mass', 'cladeBat'))
```




```{r biblio, message = FALSE}
write.bibtex(file="doc/references.bib")
```

