

This section calculates the percent differences mentioned in the discussion and abstract.

We calculated rodent and bat estimates at `log(body mass) = 0` to remove the effect
of body mass.
Since all these variables were log-transformed, we then exponentiated
both rodent and bat estimates before computing the proportional, then percentage, 
difference:

\begin{equation}
    100 \times \frac{ \exp\left( \beta_0 + \beta_{\text{clade}} \right) - 
        \exp\left( \beta_0 \right) }{ \exp\left( \beta_0 \right) }
\end{equation}


where $\beta_0$ is the intercept and $\beta_{\text{clade}}$ the coefficient for clade.


```{r results-percent-diff}
perc_diff <- function(.mod) {
    
    # Detect log-transform from formula
    logged <- grepl('^log', paste(.mod$formula)[2])
    
    coefs <- {.mod %>% summary %>% coef}[,'Estimate']
    rodents <- coefs[['(Intercept)']]
    bats <- rodents + coefs[['cladeBat']]
    if (logged) {
        rodents <- exp(rodents)
        bats <- exp(bats)
    }
    prop_diff <- (bats - rodents) / rodents
    return(prop_diff * 100)
}
```


```{r results-percent-diff-nsa}
pd_nsa <- perc_diff(plm_mods$spp$log_nsa) %>% 
    signif(digits = 4)
```
```{r results-percent-diff-sef}
pd_sef <- mean(c(
    perc_diff(plm_mods$pos$prox$log_sef),
    perc_diff(plm_mods$pos$mid$log_sef),
    perc_diff(plm_mods$pos$dist$log_sef))) %>% 
    signif(digits = 4)
```
```{r results-percent-diff-ent-dens}
pd_ed <- mean(c(
    perc_diff(plm_mods$pos$prox$log_enterocyte_density), 
    perc_diff(plm_mods$pos$mid$log_enterocyte_density), 
    perc_diff(plm_mods$pos$dist$log_enterocyte_density))) %>% 
    signif(digits = 4)
```


- __NSA\:__ $`r pd_nsa`$\%
- __SEF\:__ $`r pd_sef`$\%
- __Enterocyte density\:__ $`r pd_ed`$\%


