--- 
title: "phylolm results" 
author: "Lucas Nell" 
date: "`r format(Sys.Date(), '%d %b %Y')`" 
output: 
  github_document: 
    toc: true 
    toc_depth: 1 
editor_options: 
  chunk_output_type: console
--- 
 
```{r setup, include = FALSE, cache = FALSE} 
knitr::opts_chunk$set(echo = FALSE) 
knitr::opts_knit$set(root.dir = normalizePath("..")) 
``` 
 
 
This script displays results from 
[`phylolm`](https://doi.org/10.1093/sysbio/syu005). 
Column `estimate` is the coefficient estimate for the specified `X`, while 
`lower` and `upper` are lower and upper bounds of the 95% confidence interval
for the coefficient estimate obtained by parametric bootstrapping.
Multiple `X` parameters for a given `Y` indicate that both `X`s were included
in the regression with `Y`, *not* that separate regressions were performed for 
each `X`.
 
 
```{r libraries} 
# Packages needed for this script 
suppressPackageStartupMessages({
    library(readr)
    library(dplyr)
    library(tidyr)
    library(stringr)
    library(purrr)
}) 
``` 
 
 
```{r parse_output} 
summ_df <- read_csv('output/models_summaries.csv', col_types = 'cccddd') %>%  
    mutate(X = ifelse(X == 'taxonBat', 'taxon', X),
           pos = ifelse(is.na(pos), 'all', pos)) %>%  
    rename(estimate = value) %>%
    select(pos, Y, X, estimate, lower, upper) %>% 
    arrange(pos, Y, X) %>%
    nest(-pos)
``` 
 
 
# Positions combined 
 
```{r spp_summs} 
knitr::kable(filter(summ_df, pos == 'all')$data[[1]]) 
``` 
 
 
# Proximal 
 
```{r prox_summs} 
knitr::kable(filter(summ_df, pos == 'prox')$data[[1]])
``` 
 
 
# Medial 
 
```{r med_summs} 
knitr::kable(filter(summ_df, pos == 'med')$data[[1]])
``` 
 
 
# Distal 
 
```{r dist_summs} 
knitr::kable(filter(summ_df, pos == 'dist')$data[[1]])
``` 
 