




### Model estimates

Tables \ref{tab:results-spp-summs}--\ref{tab:results-dist-summs} summarize the 
coefficient estimates for all `phylolm` model fits. 
The "Y" column indicates the dependent variable for the model, and 
vertical whitespace separates rows of different models.
When the "parameter" column is of the form $\beta_X$, it refers to the coefficient 
for independent variable $X$. $X$ is one of the following:

- clade: binary variable whose value is 1 if the species is a bat (0 if a rodent)
- mass: log-transformed body mass
- omniv.: binary variable whose value is 1 if the species is omnivorous
- carniv.: binary variable whose value is 1 if the species is carnivorous

When the "parameter" column is of the form phylo$_Y$, it refers to the 
phylogenetic signal parameter $Y$;
$Y$ takes different values depending on the phylogenetic error term:
$\lambda$ for Pagel's lambda and $\alpha$ for Ornstein-Uhlenbeck.


```{r results-spp-summs} 
filter(summ_df, pos == 'all', 
                          model == 'phylolm')$data[[1]] %>% 
    arrange(Y, parameter) %>%
    mutate(Y = ifelse(Y == lag(Y, default = 'XXX'), "", Y) %>% clean_Y(track_log = FALSE),
           parameter = greekify(parameter)) %>% 
    select(-phy_model) %>%
    pretty_df(., caption = str_c('\\texttt{phylolm} model summaries for analyses ',
                                 'by species only. ',
                                 'All Y variables were log-transformed for these ',
                                 'analyses'))
``` 


```{r results-prox-summs} 
filter(summ_df, pos == 'prox', model == 'phylolm')$data[[1]] %>% 
    arrange(Y, phy_model, parameter) %>%
    mutate(Y = ifelse(Y == lag(Y, default = 'XXX') & 
                          phy_model == lag(phy_model, default = 'XXX'),
                      "", Y) %>% clean_Y(),
           Y = ifelse(phy_model == 'OUfixedRoot' & Y != '', 
                      str_c(Y, '$^{\\star}$'), Y),
           parameter = greekify(parameter)) %>% 
    select(-phy_model) %>% 
    pretty_df(., caption = str_c('\\texttt{phylolm} model summaries for proximal ',
                                 'intestinal segment analyses. ',
                                 '$\\dagger$ indicates Y variables that were ',
                                 'log-transformed. ',
                                 '$\\star$ indicates the model that was fit using ',
                                 'the Ornstein-Uhlenbeck model for the phylogenetic ',
                                 'error term (see ``Uncertainty in regression of ',
                                 "proximal crypt width on clade'' below for more ",
                                 'information on this model fit).'))
``` 


```{r results-med-summs} 
filter(summ_df, pos == 'med', model == 'phylolm')$data[[1]] %>% 
    arrange(Y, parameter) %>%
    mutate(Y = ifelse(Y == lag(Y, default = 'XXX'), "", Y) %>% clean_Y(),
           parameter = greekify(parameter)) %>% 
    select(-phy_model) %>% 
    pretty_df(., caption = str_c('\\texttt{phylolm} model summaries for medial ',
                                 'intestinal segment analyses. ',
                                 '$\\dagger$ indicates Y variables that were ',
                                 'log-transformed.'))
``` 


```{r results-dist-summs}
filter(summ_df, pos == 'dist', model == 'phylolm')$data[[1]] %>% 
    arrange(Y, parameter) %>%
    mutate(Y = ifelse(Y == lag(Y, default = 'XXX'), "", Y) %>% clean_Y(),
           parameter = greekify(parameter)) %>% 
    select(-phy_model) %>% 
    pretty_df(., caption = str_c('\\texttt{phylolm} model summaries for distal ',
                                 'intestinal segment analyses. ',
                                 '$\\dagger$ indicates Y variables that were ',
                                 'log-transformed.'))
``` 


### Uncertainty in regression of proximal crypt width on clade

The model for `crypt_width ~ clade` in the proximal segment, when using Pagel's 
$\lambda$, estimates a very strong phylogenetic signal 
(Table \ref{tab:results-prox-summs}).
When providing a low starting value for $\lambda$ ($\lambda = 0.1$), 
the model converges to a very weak phylogenetic signal ($\lambda = 10^{-7}$),
but it still has a higher log likelihood at the strong phylogenetic signal 
($\ell = 68.14$ at $\lambda = 0.993$) than at the weak signal 
($\ell = 66.88$ at $\lambda = 10^{-7}$).

However, this model is sensitive to starting values, which suggests multiple peaks in
the likelihood profile.
Moreover, the models for the other segments show very weak phylogenetic signal 
($\lambda = 10^{-7}$ for both), 
and this model re-run with the Ornstein-Uhlenbeck model for phylogenetic error 
("OU"; `OUfixedRoot` in `phylolm`) has a higher log likelihood and
shows a much weaker phylogenetic signal ($\ell = 68.54$ and $\alpha = 0.0156$) than
the version run with Pagel's lambda.
Thus the model likely had convergence issues using Pagel's lambda, so we
replaced the original model with one using the OU error model.




### Influential value in regression of intestinal length on clade

#### Jackknifing

We used jackknifing to determine how influential each point is to the coefficient 
estimates in the `phylolm` model for `log_intestinal_length ~ clade`.
See equation \ref{eq:jackknife} for how I calculated influence absolute Z-scores.
Figure \ref{fig:results-influence-plots}
show that one species (*Artibeus lituratus*) has a disproportionate effect on all 
three coefficient estimates.


```{r results-influence-jack, cache = FALSE}
# (See `R/model_summaries.R` for the `jack_phylolm` function.)
int_jack <- jack_phylolm(plm_mods$spp$log_intestinal_length, 
                         phy = get_tr('spp'))
```



```{r results-influence-plots, fig.height = 6.5, fig.width = 6.5, fig.cap = "Influential points for the regression of intestinal length on clade and body mass. (A) Intestinal length against body mass where point sizes are proportional to their influence absolute Z-score, color indicates clade, and both axes are on the log scale. (B) Influence absolute Z-scores by species. For both plots, facet labels indicate which parameter influence-values pertain to."}
# Adding morphometric data to the jackknife data frame
jack_df <- dplyr::left_join(int_jack, get_df('spp'), by = 'species') %>% 
    select(diet, clade, species, estimate, influence, log_mass, log_intestinal_length) %>% 
    mutate(estimate = factor(estimate, levels = c('cladeBat', 'log_mass', 'intercept'), 
                             labels = c('clade', 'body mass', 'intercept')),
           species = reorder(species, abs(influence)))
infl_p <- list()

# Plot showing context of influential values
infl_p[['A']] <- jack_df %>%
    # Exponentiate variables for plotting so we can more transparently
    # provide axis labels of the non-transformed numbers.
    # The same is done below when plotting the raw data.
    mutate_at(vars(log_mass, log_intestinal_length), exp) %>% 
    ggplot(aes(log_mass, log_intestinal_length, color = clade, size = abs(influence))) +
    geom_point() +
    # All these plots include log(mass) on the x-axis so we're including this here.
    scale_x_continuous('Body mass (g)', trans = 'log', breaks = 10 * 4^(0:2),
                       limits = exp(c(2.043877, 5.199205)), expand = c(0,0)) +
    scale_y_continuous('Intestinal length (cm)', trans = 'log', breaks = 10*2^(0:2)) +
    facet_grid(~ estimate) +
    scale_size_area('Influence\n|Z-score|:', breaks = 1:2, max_size = 4) +
    scale_color_manual('Clade:', values = c('#9ecae1', '#de2d26')) +
    theme(legend.title = element_text(size = 10, face = 'bold.italic'),
          legend.position = 'left', 
          plot.margin = margin(t = 5.5, r = 5.5, b = 0, l = 5.5))


# Plotting to show better the actual |Z-scores|
infl_p[['B']] <- jack_df %>%
    ggplot(aes(abs(influence), species, color = estimate)) +
    geom_segment(aes(yend = species), xend = 0, color = 'gray40') +
    geom_point(size = 2, color = 'dodgerblue') +
    scale_color_manual(values = c('#1b9e77','#d95f02','#7570b3')) +
    facet_wrap(~ estimate) +
    xlab("Influence |Z-score|") +
    theme(axis.title.y = element_blank(),
          axis.text.y = element_text(face = 'italic'),
          strip.text = element_blank(),
          strip.background = element_blank(),
          plot.margin = margin(t = 0, r = 5.5, b = 5.5, l = 5.5))

cowplot::plot_grid(plotlist = infl_p, ncol = 1, axis = 'lb', align = 'hv', 
                   rel_heights = c(3, 4), 
                   labels = LETTERS[1:2], label_fontface = 'plain')
```



#### Re-running model without influential point

```{r results-influence-firstrun}
influence_og <- plm_mods$spp$log_intestinal_length
first_run_df <- cbind(cbind(estimate = c(influence_og$coefficients, 
                                         lambda = influence_og$optpar)),
                      t(influence_og$bootconfint95)[c("(Intercept)", "cladeBat", 
                                                      "log_mass", "lambda"), ]) %>% 
    as_tibble %>% 
    rename(lower = `2.5%`, upper = `97.5%`) %>% 
    mutate(parameter = c("(Intercept)", "cladeBat", "log_mass", 
                         "lambda")) %>% 
    select(parameter, everything())
```


```{r results-influence-rerun, cache = TRUE}
spp_df <- get_df('spp')
spp_tr <- get_tr('spp')
.df <- get_df('spp')
.df <- .df[.df$species != 'Artibeus lituratus',]
.tr <- get_tr('spp') %>% filter_tr(.df$species)
set.seed(1018666821)
suppressWarnings(  # gives warning about low lambda
    influence_rerun <- phylolm(log_intestinal_length ~ clade + log_mass, 
                               model = 'lambda', data = .df, boot = 2000, 
                               phy = .tr))
```



Table \ref{tab:results-re-run-summ-dfs} compares summaries of this model 
with and without the influential value.
When that point is removed, we are much less sure about mass's influence on 
intestinal length.
It does not, however, change our conclusions regarding the influence of clade
(P-value for $\beta_{\text{clade}}$ with influential point:
`r sprintf("%.3f", pval(influence_og, 'cladeBat'))`;
P without it: `r sprintf("%.3f", pval(influence_rerun, 'cladeBat'))`).


```{r results-re-run-summ-dfs}
rerun_df <- cbind(cbind(estimate = c(influence_rerun$coefficients, 
                                     lambda = influence_rerun$optpar)),
                  t(influence_rerun$bootconfint95)[c("(Intercept)", "cladeBat", 
                                                     "log_mass", "lambda"), ]) %>% 
    as_tibble %>% 
    rename(lower = `2.5%`, upper = `97.5%`) %>% 
    mutate(parameter = c("(Intercept)", "cladeBat", "log_mass", 
                         "lambda")) %>% 
    select(parameter, everything())
list(first_run_df, rerun_df) %>% 
    lapply(function(x) {
        mutate(x, parameter = greekify(parameter))
    }) %>% 
    pretty_df(., caption = str_c('\\texttt{phylolm} model estimates for regression of ',
                                 'log(intestinal length) on clade and log(body mass), ',
                                 'with (left) and  without (right) influential point ',
                                 'present in dataset.'))
```

