

This section displays all results from analyses using
`phylolm::phylolm` `r citep(citation("phylolm"))` and 
`phyr::cor_phylo`
`r citep("10.1111/j.1365-2435.2009.01596.x")`.
The `phyr` package is available from <http://github.com/daijiang/phyr>.

In the tables below, the "estimate" column is the maximum likelihood estimate 
for the specified parameter, 
while columns "lower" and "upper" are lower and upper bounds of the 95% confidence 
interval for the parameter estimate.
"P" is the P-value for a parameter not being equal to zero.
Both P-values and CIs were obtained by parametric bootstrapping of the `phylolm`
or `cor_phylo` model.
We do not present P-values for phylogenetic signals because they were bound
above zero by the models; thus confidence intervals are more informative.


 
```{r results-libraries, echo = FALSE, cache = FALSE}
# Packages needed for this script 
suppressPackageStartupMessages({
    library(readr)
    library(dplyr)
    library(tidyr)
    library(stringr)
    library(purrr)
    library(phylolm)
    library(ape)
    library(ggplot2)
    library(grid)
    library(gridExtra)
    library(phyr)
})
# Set default ggplot2 theme
# setting default `ggplot2` theme
theme_set(theme_classic() %+replace% 
              theme(strip.background = element_blank(),
                    strip.text = element_text(size = 11, face = 'italic'),
                    legend.title = element_blank(),
                    legend.background = element_blank(),
                    plot.title = element_text(size = 14, hjust = 0)))
# Sourcing scripts from R directory
invisible(sapply(list.files('R', '*.R', full.names = TRUE), source))

# Function to turn parameter values to Greek versions
greekify <- function(parameters) {
    if (is.factor(parameters)) parameters <- paste(parameters)
    sapply(
            parameters, 
            function(.p){
                switch(.p, 
                       `(Intercept)` = "$\\beta_0$",
                       cladeBat = "$\\beta_{\\text{clade}}$",
                       dietOmnivorous = "$\\beta_{\\text{omniv.}}$",
                       dietProtein = "$\\beta_{\\text{carniv.}}$",
                       lambda = "phylo$_{\\lambda}$",
                       alpha = "phylo$_{\\alpha}$",
                       d = "phylo$_d$",
                       sprintf("$\\beta_{\\text{%s}}$", gsub("log_", "", .p))
                   )
            }
        )
}

# Clean up Y variables for tables
clean_Y <- function(Y, track_log = TRUE) {
    Yp <- Y %>% 
        str_replace_all('log_', '') %>% 
        str_replace_all('_', ' ') %>% 
        str_replace_all('sef', 'SEF') %>% 
        str_replace_all('nsa', 'NSA') %>% 
        str_replace_all('absorp', 'absorption') %>% 
        str_replace_all('vill ', 'villous ')
    if (track_log) {
        Ypp <- str_c(Yp, '$^{\\dagger}$')
        return(ifelse(str_detect(Y, '^log_'), Ypp, Yp))
    } else {
        return(Yp)
    }
}


# Parse scientific notation into a string for LaTeX
parse_scinot <- function(string) {
    if (str_detect(string, 'e')) {
        string <- str_split(string, 'e') %>% 
            map_chr(~ sprintf('%s \\mathrm{e}{%s}', .x[1], as.numeric(.x[2])))
    }
    return(sprintf('$%s$', string))
}


# Function to print data frames for LaTeX output using booktabs package
pretty_df <- function(df_, caption = NULL) {
    edit_nums <- function(.df){
        .df %>% 
            mutate_if(is.numeric, 
                      function(x) {
                          x %>% sprintf(fmt = '%.4g') %>% 
                              sapply(., parse_scinot)
                      })
    } 
    if (is(df_, 'list')) {
        df_ <- lapply(df_, edit_nums)
    } else {
        df_ <- edit_nums(df_)
    }
    kable_obj <- knitr::kable(df_, format = 'latex', booktabs = TRUE, caption = caption,
                              escape = FALSE, format.args = list(floating = FALSE))
    kable_str <- unlist(str_split(kable_obj, '\n'))
    kable_str <- kable_str[kable_str != "\\addlinespace"]
    for (i in 1:(length(kable_str)-1)) {
        if (grepl('^ &', kable_str[i]) & !grepl('^ &', kable_str[(i+1)]) & 
            !grepl('^\\bottomrule', kable_str[(i+1)])) {
            kable_str[i] <- str_c(kable_str[i], "\n", "\\addlinespace")
        }
    }
    new_kable_str <- str_c(kable_str, collapse = '\n')
    kable_obj[1] <- new_kable_str
    return(kable_obj)
}
``` 
 
 
```{r results-parse-output, echo = FALSE} 
summary_df <- read_csv('output/models_summaries.csv', col_types = 'ccccdddd') %>%
    mutate(X = ifelse(X == 'cladeBat', 'clade', X),
           pos = ifelse(is.na(pos), 'all', pos)) %>%  
    rename(estimate = value) %>%
    select(pos, Y, X, phy_model, estimate, lower, upper, P) %>% 
    mutate_if(is.numeric, function(s) signif(s, digits = 4)) %>% 
    mutate(model = ifelse(phy_model == 'OU', 'cor_phylo', 'phylolm'),
           P = ifelse(X %in% c('lambda', 'd', 'alpha'), "–", sprintf("$%.3f$", P)),
           # Explicitly ordering parameters:
           X = factor(X, 
                      levels = c("clade", "log_mass", 
                                 "dietOmnivorous", "dietProtein", 
                                 "log_clear", "log_total_enterocytes", 
                                 "lambda", "alpha", "d"))) %>% 
    rename(parameter = X) %>% 
    nest(-model, -pos)
plm_mods <- list(absorp = read_rds('output/models_absorp.rds'),
               diet = read_rds('output/models_diet.rds'),
               pos = read_rds('output/models_pos.rds'),
               spp = read_rds('output/models_spp.rds'))
cp_mod <- read_rds('output/models_cor_phylo.rds')
cp_mod_refits <- read_rds('output/models_cor_phylo_refits.rds')
```












