--- 
title: "phylolm results" 
author: "Lucas Nell" 
date: "`r Sys.setenv(TZ='America/Chicago'); format(Sys.Date(), '%d %b %Y')`" 
output: 
  github_document: 
    toc: true 
    toc_depth: 1 
--- 
 
```{r setup, include = FALSE, cache = FALSE} 
knitr::opts_chunk$set(echo = FALSE) 
knitr::opts_knit$set(root.dir = normalizePath(".."))
``` 
 
 
This script displays results from analyses using
[`phylolm::phylolm`](https://doi.org/10.1093/sysbio/syu005) and 
a modified version of [`ape::corphylo`](https://CRAN.R-project.org/package=ape)
(paper link 
[here](http://onlinelibrary.wiley.com/doi/10.1111/j.1365-2435.2009.01596.x/full)).

The `parameter` column indicates which variable (either an X variable or phylogenetic 
signal) that row's numbers refer to.
Column `estimate` is the parameter estimate for the specified `parameter`, while 
`lower` and `upper` are lower and upper bounds of the 95% confidence interval
for the coefficient estimate, obtained by parametric bootstrapping.
`P` is the p-value for an estimate not equal to zero; this is obtained by the same 
bootstrapping as for the CI.
I do not present P-values for phylogenetic signals because they were bound
above zero by the models.
Thus confidence intervals were more informative.

Multiple `parameter` values for a given `Y` *never* indicate that separate 
analyses were performed for each `parameter`.

> This page uses the variable name `clade` for the dummy variable indicating whether
> a species is a bat, but in the rest of the scripts in this repo, this variable 
> is referred to as `taxon`.
 
 
```{r libraries} 
# Packages needed for this script 
suppressPackageStartupMessages({
    library(readr)
    library(dplyr)
    library(tidyr)
    library(stringr)
    library(purrr)
    library(phylolm)
    library(ape)
})
devtools::load_all('corphyloCpp')
``` 
 
 
```{r parse_output} 
summ_df <- read_csv('output/models_summaries.csv', col_types = 'cccdddd') %>%
    mutate(X = ifelse(X == 'taxonBat', 'clade', X),
           pos = ifelse(is.na(pos), 'all', pos)) %>%  
    rename(estimate = value) %>%
    select(pos, Y, X, estimate, lower, upper, P) %>% 
    mutate_if(is.numeric, function(s) sprintf('%.4g', s)) %>% 
    mutate(model = ifelse((X %in% c('absorp', 'log_clear', 'd') | 
                               Y %in% c('absorp', 'log_clear')) & 
                              !X %in% c('clade', 'lambda', 'log_mass'), 
                          'corphylo','phylolm'),
           P = ifelse(X %in% c('lambda', 'd'), "–", P)) %>% 
    rename(parameter = X) %>% 
    nest(-model, -pos)
plm_mods <- list(absorp = read_rds('output/models_absorp.rds'),
               diet = read_rds('output/models_diet.rds'),
               pos = read_rds('output/models_pos.rds'),
               spp = read_rds('output/models_spp.rds'))
cp_mods <- read_rds('output/models_corphylo.rds')
``` 
 

# `phylolm` estimates

Below, the `parameter` column names coefficients (i.e., X variables), except for
when `parameter == "lambda"`.
When `parameter == "lambda"`, this refers to the phylogenetic correlation parameter
(Pagel's lambda).


## Positions combined 
 
```{r spp_summs} 
knitr::kable(filter(summ_df, pos == 'all', model == 'phylolm')$data[[1]] %>% 
    arrange(Y, parameter) %>%
    mutate(Y = ifelse(Y == lag(Y, default = 'XXX'), "", Y)))
``` 
 
 
## Proximal 
 
```{r prox_summs} 
knitr::kable(filter(summ_df, pos == 'prox', model == 'phylolm')$data[[1]] %>% 
    arrange(Y, parameter) %>%
    mutate(Y = ifelse(Y == lag(Y, default = 'XXX'), "", Y)))
``` 
 
 
## Medial 
 
```{r med_summs} 
knitr::kable(filter(summ_df, pos == 'med', model == 'phylolm')$data[[1]] %>% 
    arrange(Y, parameter) %>%
    mutate(Y = ifelse(Y == lag(Y, default = 'XXX'), "", Y)))
``` 
 
 
## Distal 
 
```{r dist_summs} 
knitr::kable(filter(summ_df, pos == 'dist', model == 'phylolm')$data[[1]] %>% 
    arrange(Y, parameter) %>%
    mutate(Y = ifelse(Y == lag(Y, default = 'XXX'), "", Y)))
``` 


# `corphylo` estimates

Below, when `parameter != "d"`, that row refers to the correlation between the
`Y` and `parameter` variables. There are no real X and Y variables here.

When `parameter == "d"`, this row refers to the phylogenetic signal from an 
Ornstein-Uhlenbeck process for variable `Y`.

# `Clearance` and `SEF`

```{r clear_sef_kable}
knitr::kable(filter(summ_df, model == 'corphylo')$data[[1]][1:3,])
```

# `Clearance` and `log_enterocyte_density`

```{r clear_ed_kable}
knitr::kable(filter(summ_df, model == 'corphylo')$data[[1]][4:6,])
```


# `Absorption` and `log_total_enterocytes`

```{r absorp_te_kable}
knitr::kable(filter(summ_df, model == 'corphylo')$data[[1]][7:9,])
```




# Percent differences

This is for the percent differences in the discussion and abstract.

```{r percent_diff, echo = TRUE}
perc_diff <- function(.m, log_trans = FALSE) {
    .c <- {.m %>% summary %>% coef}[,'Estimate']
    .r <- .c[['(Intercept)']]
    .b <- .r + .c[['taxonBat']]
    .d <- .b - .r
    .d / .r
    if (log_trans) {
        prop_diff <- {exp(.c[2] + .c[1]) - exp(.c[1])} / exp(.c[1])
    } else {
        prop_diff <- .c[['taxonBat']] / .c[['(Intercept)']]
    }
    return(prop_diff * 100)
}

.m <- plm_mods$spp$nsa


# here include jackknife evidence that row 1 is super influential and re-run model 
# without it (present those results in the paper too)

mm <- phylolm(log(sef) ~ taxon + log_mass, get_df('spp'), model = 'lambda', 
              phy = get_tr('spp'))
summary(mm)

nd <- as_tibble(expand.grid(log_mass = seq(2, 5, length.out = 100), taxon = c('Bat', 'Rodent'))) %>% 
    mutate(taxon = factor(taxon, levels = c('Rodent', 'Bat')))
nd$intestinal_length <- exp(as.numeric(predict(mm, newdata = nd)))
nd$sef <- exp(as.numeric(predict(mm, newdata = nd)))
# nd$nsa <- as.numeric(predict(mm, newdata = nd))^2
nd$nsa2 <- as.numeric(predict(.m, newdata = nd))


library(ggplot2)
theme_set(theme_bw())
source(".Rprofile")

p_df <- lapply(c('prox', 'med', 'dist'), function(p) get_df('pos', p) %>% mutate(pos = p)) %>% 
    bind_rows %>% 
    as_tibble %>% 
    mutate(pos = factor(pos, levels = c('prox', 'med', 'dist')))

p_df %>% 
    ggplot(aes(log_mass, log_enterocyte_density, color = taxon)) +
    geom_point() +
    facet_grid(. ~ pos)

invisible(sapply(list.files('R', '*.R', full.names = TRUE), source))


get_df('spp') %>% 
    ggplot(aes(log_mass, log(intestinal_length), color = taxon)) +
    geom_point()

source("R/get_data.R")
plm <- plm_mods$spp$intestinal_length
phy = get_tr('spp')




# Function to do jackknifing on phylolm object to find influential points (i.e., species)
jack_phylolm <- function(plm, phy, return_fxn = NULL) {
    
    if (is.null(return_fxn)) {
        return_fxn <- function(.plm) return(t(coef(.plm)))
    }
    
    X_names <- gsub("\\s+", "", unlist(strsplit(paste(plm$formula)[3], '\\+|\\*')))
    
    df_ <- cbind(plm$X[,-1], as.data.frame(plm$y))
    colnames(df_)[length(colnames(df_))] <- paste(plm$formula)[2]
    # Dealing with factors
    for (xn in X_names) {
        ii <- which(grepl(xn, colnames(df_)))
        # This one must be taxon 
        if (colnames(df_)[ii][1] == 'taxonBat') {
            new_col <- factor(ifelse(df_$taxonBat == 1, 'Bat', 'Rodent'),
                              levels = c('Rodent', 'Bat'))
            df_[,xn] <- new_col
            df_ <- df_[,-ii]
        } else if (length(ii) == 2) {
        # This one must be diet
            new_col <- factor(ifelse(df_$dietOmnivorous == 1, 'Omnivorous', 
                                     ifelse(df_$dietProtein == 1, 'Protein', 'Herbivorous')),
                              levels = c('Herbivorous', 'Omnivorous', 'Protein'))
            df_[,xn] <- new_col
            df_ <- df_[,-ii]
        }
    }; rm(xn, ii, new_col)
    
    comp_ <- return_fxn(plm)
    
    jack_df <- lapply(1:nrow(df_), 
           function(i) {
               .df <- df_[-i,]
               .tr <- ape::drop.tip(phy, phy$tip.label[!phy$tip.label %in% rownames(.df)])
               suppressWarnings(
                   .plm <- update(plm, data = .df, phy = .tr, boot = 0)
               )
               return(return_fxn(.plm) - comp_)
           }) %>% 
        do.call(what = rbind) %>% 
    as_tibble %>% 
    rename(intercept = `(Intercept)`) %>% 
    mutate(species = rownames(df_)) %>% 
    gather('estimate', 'influence', -species)
    
    return(jack_df)
}






spp_df <- get_df('spp')
spp_tr <- get_tr('spp')



int_jack <- jack_phylolm(plm_mods$spp$intestinal_length, get_tr('spp'))

int_jack %>% 
    ggplot(aes(abs(influence), reorder(species, abs(influence)), color = estimate)) +
    geom_point() +
    xlab('Jackknife influence value') +
    theme(axis.title.y = element_blank(), axis.text.y = element_text(face = 'italic'),
          legend.position = c(0.9, 0.1), legend.justification = c(1, 0),
          legend.background = element_rect(color = 'black', size = 0.25))





which(spp_df$species == 'Artibeus lituratus')
which(spp_df$species == 'Euryoryzomys russatus')
spp_df[c(1, 12),]



spp_df %>% 
    ggplot(aes(log_mass, log_intestinal_length, color = taxon)) +
    geom_point()

.df <- .df[.df$species != 'Artibeus lituratus',]
.tr <- get_tr('spp') %>% filter_tr(spp[spp != 'Artibeus lituratus'])
z <- phylolm(intestinal_length ~ taxon + log_mass, model = 'lambda',
        data = .df, boot = 2000, phy = .tr)
z %>% summary
pval(z, c('taxonBat', 'log_mass'))

summary(plm_mods$spp$intestinal_length)
pval(plm_mods$spp$intestinal_length, c('taxonBat', 'log_mass'))


hist(resid(plm_mods$pos$prox$log_enterocyte_density))
hist(resid(mm))

```

## NSA corrected for mass
```{r percent_diff_nsa, echo = TRUE}
cat(sprintf('%.4g %%\n', perc_diff(plm_mods$spp$nsa)))
```

## SEF
```{r percent_diff_sef, echo = TRUE}
cat(sprintf('%.4g %%\n', mean(c(
    perc_diff(plm_mods$pos$prox$sef),
    perc_diff(plm_mods$pos$med$sef),
    perc_diff(plm_mods$pos$dist$sef)
))))
```

## Enterocyte density
```{r percent_diff_ent_dens, echo = TRUE}
cat(sprintf('%.4g %%\n', mean(c(
    perc_diff(plm_mods$pos$prox$log_enterocyte_density, TRUE), 
    perc_diff(plm_mods$pos$med$log_enterocyte_density, TRUE), 
    perc_diff(plm_mods$pos$dist$log_enterocyte_density, TRUE)
))))
```





# SEF on Taxon and log(Mass)

This is to determine whether there's an effect of body mass on SEF. It appears there
is not.

```{r sef_taxon_mass, echo = TRUE}
invisible(sapply(list.files('R', '*.R', full.names = TRUE), source))
tr <- get_tr('spp')
spp_df <- get_df('spp') %>% rename(clade = taxon)
set.seed(940318092)
mod <- suppressWarnings(
    phylolm(log_sef ~ clade + log_mass, data = spp_df, phy = tr, model = "lambda",
            boot = 2000))
# P-values for coefficient estimates != 0, using parametric bootstrapping:
pval(mod, c('log_mass', 'cladeBat'))
```

