--- 
title: "phylolm results" 
author: "Lucas Nell" 
date: "`r Sys.setenv(TZ='America/Chicago'); format(Sys.Date(), '%d %b %Y')`" 
output: 
  github_document: 
    toc: true 
    toc_depth: 1 
editor_options: 
  chunk_output_type: console
--- 
 
```{r setup, include = FALSE, cache = FALSE} 
knitr::opts_chunk$set(echo = FALSE) 
knitr::opts_knit$set(root.dir = normalizePath("..")) 
``` 
 
 
This script displays results from 
[`phylolm`](https://doi.org/10.1093/sysbio/syu005). 
Column `estimate` is the coefficient estimate for the specified `X`, while 
`lower` and `upper` are lower and upper bounds of the 95% confidence interval
for the coefficient estimate obtained by parametric bootstrapping.
Multiple `X` parameters for a given `Y` indicate that both `X`s were included
in the regression with `Y`, *not* that separate regressions were performed for 
each `X`.
 
 
```{r libraries} 
# Packages needed for this script 
suppressPackageStartupMessages({
    library(readr)
    library(dplyr)
    library(tidyr)
    library(stringr)
    library(purrr)
}) 
``` 
 
 
```{r parse_output} 
summ_df <- read_csv('output/models_summaries.csv', col_types = 'cccddd') %>%  
    mutate(X = ifelse(X == 'taxonBat', 'clade', X),
           pos = ifelse(is.na(pos), 'all', pos)) %>%  
    rename(estimate = value) %>%
    select(pos, Y, X, estimate, lower, upper) %>% 
    arrange(pos, Y, X) %>%
    nest(-pos)
``` 
 
 
# Positions combined 
 
```{r spp_summs} 
knitr::kable(filter(summ_df, pos == 'all')$data[[1]]) 
``` 
 
 
# Proximal 
 
```{r prox_summs} 
knitr::kable(filter(summ_df, pos == 'prox')$data[[1]])
``` 
 
 
# Medial 
 
```{r med_summs} 
knitr::kable(filter(summ_df, pos == 'med')$data[[1]])
``` 
 
 
# Distal 
 
```{r dist_summs} 
knitr::kable(filter(summ_df, pos == 'dist')$data[[1]])
``` 

# SEF on Taxon and log(Mass)

This is to determine whether there's an effect of body mass on SEF. It appears there 
is not.

```{r sef_taxon_mass, echo = TRUE}
suppressPackageStartupMessages(library(phylolm))
invisible(sapply(list.files('R', '*.R', full.names = TRUE), source))
tr <- get_tr('spp')
spp_df <- get_df('spp') %>% rename(clade = taxon)
set.seed(940318092)
mod <- suppressWarnings(
    phylolm(sef ~ clade + log_mass, data = spp_df, phy = tr, model = "lambda", 
            boot = 2000))
ci(mod, 'log_mass')
```



# Percent differences

This is for the abstract.

```{r percent_diff, echo = TRUE}
models <- list(absorp = read_rds('output/models_absorp.rds'),
               pos = read_rds('output/models_pos.rds'),
               spp = read_rds('output/models_spp.rds'))
perc_diff <- function(.m, log_trans = FALSE) {
    .c <- as.numeric({.m %>% summary %>% coef}[,'Estimate'])
    if (!log_trans) return(.c[2] / .c[1])
    {exp(.c[2] + .c[1]) - exp(.c[1])} / exp(.c[1])
}

# NSA corrected for mass
perc_diff(models$spp$nsa_mass)

# SEF
mean(c(perc_diff(models$pos$prox$sef), perc_diff(models$pos$med$sef),
       perc_diff(models$pos$dist$sef)))

# Enterocyte density
mean(c(perc_diff(models$pos$prox$log_enterocyte_density, TRUE), 
       perc_diff(models$pos$med$log_enterocyte_density, TRUE), 
       perc_diff(models$pos$dist$log_enterocyte_density, TRUE)))
```

