---
title: "Rodent phylogenetic tree"
author: "Lucas Nell"
date: "`r format(Sys.Date())`"
output: github_document
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(comment = '', prompt = FALSE)
```


# Background

The mammal paper Antonio referenced
[(Bininda-Emonds et al. 2007)](http://dx.doi.org/10.1038/nature05634)
has 6 of the 10 focal rodent species.

Another paper 
[(Fabre et al. 2012)](http://dx.doi.org/10.1186/1471-2148-12-88)
has all of them, but this tree appears quite different from the others (branch tips
do not extend to the same location horizontally; I believe this is because it is based on 
genetic distances rather than time). Some species are also located on nodes, rather
than branch tips (see plots below for a visualization), and some edge lengths are missing.
I'm not sure how to resolve any of these issues.

A last paper
[(Schenk et al. 2013)](http://sysbio.oxfordjournals.org/content/62/6/837)
has 6 of the 10 rodents, as well, but they don't entirely overlap with those from 
the mammal paper.

*Akodon montensis* and *Euryoryzomys russatus* are the two species present in neither
the Bininda-Emonds nor the Schenk trees.



# Preamble

I first need to `source` the `'tree_preamble.R'` file to load necessary R packages and 
make the data frame of species names.

```{r, message = FALSE, warning = FALSE}
source('tree_preamble.R')
```



# Schenk tree

The OTOL link for this paper is
[here](https://tree.opentreeoflife.org/curator/study/view/pg_2859).

```{r}
schenk <- get_study(study_id = 'pg_2859', object_format = 'phylo')
schenk$tip.label <- gsub('_', ' ', schenk$tip.label)
schenk
```



# Fabre tree

## Downloading tree

Using `rotl` directly didn't work for this paper's tree, so I downloaded the Newick 
version directly from the link on 
[OTOL's site](https://tree.opentreeoflife.org/curator/study/view/pg_2688/?tab=metadata).
The below code chunk is in bash (i.e., via Terminal in macOS).

```{r, engine = 'bash', eval = FALSE}
wget "https://api.opentreeoflife.org/v3/study/pg_2688.tre"
mv pg_2688.tre rodents.tre
```

I then opened this file in a text editor and removed all single quotes and replaced all
spaces with underscores. This allows R to read this file appropriately. After it's read,
I can replace underscores with spaces.

```{r}
rod_tr <- read.tree('rodents.tre')
rod_tr$tip.label <- gsub('_', ' ', rod_tr$tip.label)
rod_tr$node.label <- gsub('_', ' ', rod_tr$node.label)
```

## Attempting to correct for missing edge lengths

I set edge lengths of `NA` to zero, then used `di2multi` to collapse these into
multichotomies.

```{r}
rod_tr$edge.length[is.na(rod_tr$edge.length)] <- 0
rod_tr <- di2multi(rod_tr) %>% 
    # Also removing unnecessary tips
    drop.tip(., tip = c(2260:2165, 685, 686))
rod_tr
```


## Drawing tree

If you want to make a pdf of this tree that's (somewhat) readable:
```{r, eval = FALSE}
font_size <- 0.05
pdf('rod_tree.pdf', width = 6, height = 24)
    plot(rod_tr, cex = font_size, edge.width = 0.25, no.margin = TRUE, label.offset = 0.5)
    nodelabels(frame = 'none', cex = font_size, adj = c(1, 0))
    tiplabels(frame = 'none', cex = font_size, adj = c(0, 0.5))
dev.off()
```



## Checking species presence in tree

Three species initially don't seem to be present in either node or tip labels.

```{r}
sp_df$species[sp_df$type == 'rodent' & 
                  !sp_df$species %in% c(rod_tr$node.label, rod_tr$tip.label)]
```


However, looking these up in the OTOL database, I found the following:

- *Olygoryzomys nigripes* is spelled *Oligoryzomys nigripes*
  ([link](https://tree.opentreeoflife.org/taxonomy/browse?id=752853))
- *Sooretamys angouya* is synonymous with *Oryzomys angouya*
  ([link](https://tree.opentreeoflife.org/taxonomy/browse?id=1039661))
- *Euryoryzomys russatus* is synonymous with *Oryzomys russatus*
  ([link](https://tree.opentreeoflife.org/taxonomy/browse?id=739))

When these are taken into account, all species are found in this tree.


## Species names as node labels

The following plots show the three focal species used as node rather than tip labels.

```{r}
extract.clade(rod_tr, node = 'Mus musculus') %>% 
    plot.phylo(., show.node.label = TRUE)
extract.clade(rod_tr, node = 'Thaptomys nigrita') %>% 
    plot.phylo(., show.node.label = TRUE)
extract.clade(rod_tr, node = 'Delomys sublineatus') %>% 
    plot.phylo(., show.node.label = TRUE)
```




